---
title: "Ins_comp_2_rarefaction"
author: "laura"
date: "2024-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at difff life stages.


This script generates rarefaction curves and rarefies the data, removing any samples with read depths below the thresholds.


# Install packages

```{r}

# phyloseq itself
#if(!requireNamespace("BiocManager")){
#  install.packages("BiocManager")
 #  }
#BiocManager::install("phyloseq")

# qiime2r
# to be able to convert my qiime2 files
#if (!requireNamespace("devtools", quietly = TRUE)){install.packages("devtools")}
# devtools::install_github("jbisanz/qiime2R") # current version is 0.99.20
 
# if (!requireNamespace("BiocManager", quietly=TRUE))
 #   install.packages("BiocManager")
## BiocManager::install("BiocUpgrade") ## you may need this
#BiocManager::install("MicrobiotaProcess")
 

library(phyloseq)
library(qiime2R)
library(ggplot2)

library(MicrobiotaProcess)
library(patchwork)

```


# import data

```{r}
Laura.ps = readRDS("physeq_cont_removed")
```
This is currently 20065 ASVs.
 
First, have a reminder of sample sizes

```{r}
df <- as.data.frame(sample_data(Laura.ps)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(Laura.ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_Type2))  + geom_point(alpha=0.7, size=1)+
scale_color_manual(values=c("#006347", "#97D0A7", "#7AA8D7", "#56187D", "#AC0123", "#EF9F26", "#E6CD69"))
       #alt #88DAf6, #ABA2D2
```

The lowest sample is 3974 reads (from the plot and the df)
Try rarefying to that depth and checking rarecurve. 

```{r}
ps_3974 <- rarefy_even_depth(Laura.ps, sample.size = 3974,
  rngseed = TRUE, replace = FALSE, trimOTUs = TRUE, verbose = TRUE)
```

12700 ASVs removed as they are no longer present in the dataset, we now have 7365 ASVs

Calculate the data points

```{r warning=FALSE}

# for reproducibly random number
set.seed(1024)
rareres <- get_rarecurve(obj=ps_3974, chunks=40)
```

plot 

```{r warning=FALSE}

p_rare <- ggrarecurve(obj=rareres, factorNames="Sample_Type2",
                      indexNames=c("Observe","Shannon","Simpson"),
                      ) +
          theme(legend.spacing.y=unit(0.01,"cm"),
                legend.text=element_text(size=4))

prare1 <- ggrarecurve(obj=rareres, factorNames="Sample_Type2",
                      indexNames=c("Observe", "Shannon", "Simpson")
                      ) +
          scale_fill_manual(values=c("#8DA0CB", "#E78AC3", "#E5C494", "#fb8072", "#66C2A5", "#ABA2D2", "#AEDCDA"))+
          scale_color_manual(values=c("#8DA0CB", "#E78AC3", "#E5C494", "#fb8072", "#66C2A5", "#ABA2D2", "#AEDCDA"))+
          theme_bw()+
          theme(axis.text=element_text(size=8), panel.grid=element_blank(),
                strip.background = element_rect(colour=NA,fill="grey"),
                strip.text.x = element_text(face="bold"))          

prare2 <- ggrarecurve(obj=rareres,
                      factorNames="Sample_Type2",
                      shadow=FALSE,
                      indexNames=c("Observe", "Shannon", "Simpson")
                      ) +
          scale_color_manual(values=c("#8DA0CB", "#E78AC3", "#E5C494", "#fb8072", "#66C2A5", "#ABA2D2", "#AEDCDA"))+
          theme_bw()+
          theme(axis.text=element_text(size=8), panel.grid=element_blank(),
                strip.background = element_rect(colour=NA,fill="grey"),
                strip.text.x = element_text(face="bold"))
p_rare / prare1 / prare2


```


```{r warning=FALSE}

   

prare3 <- ggrarecurve(obj=rareres,
                      factorNames="Sample_Type2",
                      shadow=FALSE,
                      indexNames=c("Observe", "Shannon")
                      ) +
          scale_color_manual(values=c("#006347", "#97D0A7", "#7AA8D7", "#56187D", "#AC0123", "#EF9F26", "#E6CD69"))+
          theme_bw()+
          theme(axis.text=element_text(size=8), panel.grid=element_blank(),
                strip.background = element_rect(colour=NA,fill="grey"),
                strip.text.x = element_text(face="bold"))+
  xlim(0,4000)


prare3


```

# and now trying a different way

```{r warning=FALSE}
library(ranacapa)
p <- ggrare(ps_3974, step = 10, color = "Sample_Type2", label = NULL, se = FALSE) + 
   theme_bw() + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  facet_wrap(~Sample_Type2)+
          scale_color_manual(values=c("#56187D", "#AC0123", "#006347", "#97D0A7", "#EF9F26", "#7AA8D7"))
 

p
```

# and now trying a different way

```{r warning=FALSE}
#library(ranacapa)
p <- ggrare(ps_3974, step = 10, color = "Sample_Type2", label = NULL, se = FALSE) + 
 #  theme_bw() + 
 # theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  facet_wrap(~Sample_Type2)+
          scale_color_manual(values=c("#56187D", "#AC0123", "#006347", "#97D0A7", "#EF9F26", "#7AA8D7"))
 

p
```


# save new rarefied phyloseq object

```{r}
saveRDS(ps_3974, "physeq_obj_3974")

```