---
title: "Ins_comp_b4-taxonomy"
author: "laura"
date: "2024-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.


This script looks at what bacterial taxa are present in different groups and sample types (big picture) will follow up after with DESeq2 and heatmaps to identify ASVs associated with different insectaries

This is from the following:
https://rpubs.com/lgschaerer/1006964 
https://stackoverflow.com/questions/72569157/r-calculate-most-abundant-taxa-using-phyloseq-object


# Install packages

```{r}
# phyloseq itself
#if(!requireNamespace("BiocManager")){
#  install.packages("BiocManager")
#   }
#BiocManager::install("phyloseq")

library(phyloseq)
library(ggplot2)
library(tidyverse)
library(dplyr) # for select command

```

# starting with the phyloseq obj

```{r}
rm(list = ls()) # clear env

ps = readRDS("physeq_obj_3974") # currently phyloseq obj

head(sample_data(ps))

```

Here I identify the top 20 genera across the whole dataset and plot those, regardless of how much is in any one sampletype and then group everything else as other.

STEP 1: identify list of top 20 genera in dataset

```{r, eval=FALSE, warning=FALSE}

T <- ps %>% 
    tax_glom(., "Genus") %>% 
    transform_sample_counts(function(x)100* x / sum(x)) %>% psmelt() %>% 
    arrange(OTU) %>% rename(OTUsID = OTU) %>% 
    select(OTUsID, Genus, SampleID, Abundance) %>%
    spread(SampleID, Abundance)

T$Mean <- rowMeans(T[, c(3:ncol(T))])

GEN <- T[, c("Genus", "Mean" ) ]

#order data frame  
GEN <- GEN[order(dplyr::desc(GEN$Mean)),]
rownames(GEN) <- NULL

head(GEN)

# subset to top 20

GEN20 <- top_n(GEN,20,Mean) # didn't need to reorder the dataframe before doing this

# make a vector for the genera of interest
GEN20 <- GEN20$Genus


```

now subset the data for plotting to contain only ASVs with these genera

This is a long way round, but to filter full dataset for these top genera, 

STEP 2: convert phloseq obj into a dataframe showing genus level abundance data (similar to above but containing slightly different info)

```{r}
genusabundance <- ps %>%
  tax_glom(taxrank = "Genus") %>%                        # Set to smallest taxonomic level you are interested in
  transform_sample_counts(function(x) {x/sum(x)} ) %>%   # Transform to rel. abundance
  psmelt()                                               # Melt to long format
head(genusabundance)
```

STEP 3: Filter, group and modify data to prepare for plotting.

select() function: Choose which variables to include in the plot. Select ALL variables that will be used for x-axis and/or facet labels in addition to all taxonomic levels you are interested in.

filter() function: Filter out all taxa with zero percent abundance. Filter can also be used to remove treatments or conditions that are not to be included in the plot.

mutate() function: Convert the taxonomic level columns to character vectors (they are factors by default), can also be used to add additional columns or modify existing columns as desired.


```{r}
all <- genusabundance %>%
  select(Phylum, Class, Family, Genus, SampleID, Abundance, Sample_Type2, Building, for_ordering) %>%
  filter(Abundance != 0) %>%
  mutate(
    Phylum = as.character(Phylum),
    Class = as.character(Class),
    Family = as.character(Family),
    Genus = as.character(Genus))
head(all)

```


STEP 4: modify the dataframe to replace all genus names that aren't the 20 of interest to '_Other' (added the underscore just as a simple way to put Other at the top of the plots)

```{r}

genus_new <- all
genus_new <- genus_new %>%
    mutate(Genus = case_when(Genus == 'Asaia' ~ 'Asaia',
                           Genus == 'Elizabethkingia' ~ 'Elizabethkingia',
                           Genus == 'Delftia' ~ 'Delftia',
                           Genus == 'Massilia' ~ 'Massilia',
                           Genus == 'Vibrio' ~ 'Vibrio',
                           Genus == 'Burkholderia-Caballeronia-Paraburkholderia' ~ 'Burkholderia-Caballeronia-Paraburkholderia',
                           Genus == 'Sphingobacterium' ~ 'Sphingobacterium',
                           Genus == 'Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium' ~ 'Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium',
                           Genus == 'Arthrospira_PCC-7345' ~ 'Arthrospira_PCC-7345',
                           Genus == 'Solitalea' ~ 'Solitalea',
                           Genus == 'Flavobacterium' ~ 'Flavobacterium',
                           Genus == 'Peredibacter' ~ 'Peredibacter',
                           Genus == 'Microbacterium' ~ 'Microbacterium',
                           Genus == 'Bifidobacterium' ~ 'Bifidobacterium',
                           Genus == 'Phreatobacter' ~ 'Phreatobacter',
                           Genus == 'Stenotrophomonas' ~ 'Stenotrophomonas',
                           Genus == 'Pedobacter' ~ 'Pedobacter',
                           Genus == 'Acinetobacter' ~ 'Acinetobacter',
                           Genus == 'Sphingomonas' ~ 'Sphingomonas',
                           Genus == 'Pseudomonas' ~ 'Pseudomonas',
                           TRUE ~ '_Other'))


length(unique(genus_new$Genus)) # just check there's 20 + 'other'


```

STEP 5: prepare the data for plotting with building and sampletype groupings

```{r}
genus_new_group <- genus_new %>%
  select(Sample_Type2, Building, Genus, Abundance, SampleID) %>%
  group_by(Sample_Type2, Building) %>%
  mutate(totalSum = sum(Abundance)) %>%
  ungroup() %>%
  group_by(Sample_Type2, Building, Genus, totalSum) %>%
  summarise(
    Abundance = sum(Abundance),
    RelAb = Abundance/totalSum) %>%
  unique()

```

STEP 6: set colours

```{r}
colours <- c(
  "_Other" = "black",
  "Acinetobacter" = "#638B27",
  "Asaia" = "#378B53", 
  "Burkholderia-Caballeronia-Paraburkholderia" = "#AC0123", 
  "Stenotrophomonas" = "#7FC3CE", 
  "Massilia" = "#97D0A7", 
  "Phreatobacter" = "#56187D", 
  "Sphingobacterium" = "#006347", 
  "Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium" = "#00008B",
  "Bifidobacterium" = "#CA5054",
  "Pedobacter" = "#DFE2A9",
  "Elizabethkingia" = "#FFC44B",
  "Microbacterium" = "#EA8F97",
  "Pseudomonas" = "#DA6610",
  "Vibrio" = "#E49DB8",
  "Arthrospira_PCC-7345" = "#9C58A1",
  "Sphingomonas" = "#00A087FF",
  "Delftia" =  "#485AB1",
  "Flavobacterium" = "#FFE8A1",
  "Peredibacter" =  "#3C5488FF",
  "Solitalea" = "#EF9F26"
)
 #     "#BA2F1E"  "#FA9E01", ""   "#A0B360"  "" "#0B8B7F" "#36A3A7", "#91B5D7" , "#E6CD69", "#7AA8D7", 


```

STEP 7: plot

```{r}
plotA <- abund_plot <- ggplot(genus_new_group)+
  geom_col(mapping = aes(x = Building, y = RelAb, fill = Genus), position = "stack", show.legend = TRUE)+
  facet_grid(~factor(Sample_Type2, levels=c("Water", "FishFood", "LarvalWater", "Larva", "Sugar", "AdultFemale")), scales = "free", space = "free")+
  ylab("Proportion of Community") +
  xlab(NULL)+
  scale_fill_manual(values = colours) +
#  theme_linedraw()+
  theme(axis.text.y = element_text(size = 14, color = "black"),
        axis.title.y = element_text(size = 14, color = "black"),
        axis.text.x = element_text(size = 14, angle = 45, vjust = 1, hjust = 1, color = "black"),
        legend.text = element_text(size = 14),
        legend.position = "right",
        legend.spacing.x = unit(0.1, 'mm'),
        legend.spacing.y = unit(0.05, 'mm'),
        plot.margin=grid::unit(c(0.3,0.3,0.3,0.3), "mm"),
        strip.text = element_text(size = 16, face = "bold", angle = 0),
        legend.title = element_text(face="bold", size = 14))+
  guides(fill=guide_legend(ncol=1,byrow=TRUE))

plotA

```
This is the one.

# now for as individual samples


```{r}
# add a column of insectary + sampleIDs so the order of the x axis is by insectary 
genus_ordered <- genus_new 
genus_ordered$ins_samp <- paste(genus_ordered$Building, genus_ordered$SampleID, sep = "-")

plotB <- abund_plot <- ggplot(genus_ordered)+
  geom_col(mapping = aes(x = ins_samp, y = Abundance, fill = Genus), position = "stack", show.legend = TRUE)+
  ylab("Proportion of Community") +
  xlab(NULL)+
  scale_fill_manual(values = colours) +
#  theme_linedraw()+
  theme(axis.text.y = element_text(size = 11, color = "black"),
        axis.title.y = element_text(size = 11, color = "black"),
        axis.text.x = element_text(size = 5, angle = 45, vjust = 1, hjust = 1, color = "black"),
        legend.text = element_text(size = 11),
        legend.position = "right",
        legend.spacing.x = unit(0.1, 'mm'),
        legend.spacing.y = unit(0.05, 'mm'),
        plot.margin=grid::unit(c(0.3,0.3,0.3,0.3), "mm"),
        strip.text = element_text(size = 16, face = "bold", angle = 0),
        legend.title = element_text(face="bold", size = 14))+
        facet_grid(~factor(Sample_Type2, levels=c("Water", "FishFood", "LarvalWater", "Larva", "Sugar", "AdultFemale")), scales = "free", space = "free")+
        guides(fill=guide_legend(ncol=1,byrow=TRUE))

plotB 


```

```{r}

```
