---
title: "Ins_comp_3c-beta_diversity_metrics"
author: "laura"
date: "2024-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.

Analysis of beta diversity - bray curtis and unweighted unifrac

https://bioconductor.org/packages/release/bioc/vignettes/MicrobiotaProcess/inst/doc//MicrobiotaProcess.html
can also look at https://www.yanh.org/2021/01/01/microbiome-r/

### Install packages

```{r}
# phyloseq itself
#if(!requireNamespace("BiocManager")){
#  install.packages("BiocManager")
#   }
#BiocManager::install("phyloseq")

# if (!requireNamespace("BiocManager", quietly=TRUE))
 #   install.packages("BiocManager")
## BiocManager::install("BiocUpgrade") ## you may need this
#BiocManager::install("MicrobiotaProcess")
 
library(phyloseq)
library(ggplot2)
library(dplyr) # for select command
#library(MicrobiotaProcess)
#library(gghalves)
#library(ggh4x)
library(ggpubr)
library(dunn.test)
#library(scales) # for the ordination plotting
library(pairwiseAdonis)
#library(gridExtra)

```

### import data

```{r}
ps = readRDS("physeq_obj_3974") # currently phyloseq obj

```

### the colours I'll use

```{r}
colours <- c(
  "AdultFemale" = "#56187D", 
  "FishFood" = "#AC0123", 
  "Larva" = "#006347", 
  "LarvalWater" = "#97D0A7", 
  "Sugar" = "#EF9F26", 
  "Sugar2" = "#E6CD69", 
  "Water" = "#7AA8D7"
)


colours2 <- c(
  "C-ACC" = "#00A087FF",
  "A-CTID" = "#3C5488FF",
  "B-MK" = "#9C58A1"
)

```

## beta diversity

Not doing Hellinger transformation (as is suggested by the microbiotaProcess tutorial/vignette) as the input data are rarefied

# bray curtis first

### firstly, all data together

description from : https://scienceparkstudygroup.github.io/microbiome-lesson/06-beta-diversity/index.html

Statistical test
To test whether the groups are different with respect to centroid and dispersion, a PERMANOVA statistical test will be performed. For this a multivariate extension of ANOVA will be used, as there are many OTU that will be used in the test. The extension is based on distances between samples. The test compares distances of samples within the same group to distances of samples from different groups. If the distance between samples from the different groups is much larger than samples from the same group, we conclude that the groups are not equal.

In order to test the significance of the result, a permutation test is used. Thus all samples are randomly mixed over the groups and the test is repeated many times. If the ratio (between group distance / within group distance) is much larger for the original data than for the permutations, we conclude there is a statistical significant difference.

The test can be applied in combination with any distance measure. 


```{r}
bray_dist = phyloseq::distance(ps, method="bray", weighted=F) # calculate distances
ordination = ordinate(ps, method="PCoA", distance=bray_dist) # ordinate

```

```{r}
# plot

brayordplot <- plot_ordination(ps, ordination, color = "Sample_Type2") #basic plot

p2 <- brayordplot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Sample_Type2))+
  scale_fill_manual(values = colours)+
  scale_color_manual(values = colours)+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggtitle("bray-curtis")


p2$layers <- p2$layers[-1] # this removed a first layer that was a solid point which had resulted in double overlaying points

p2  
```
overall differences by sample type and/or building. 

First, easier if we take the data out of the phyloseq object


```{r}
sample_data <- as.data.frame(sample_data(ps))
bray_dist_matrx = as.matrix(bray_dist)
bray_dist_df <- cbind(rownames(bray_dist_matrx), data.frame(bray_dist_matrx, row.names =NULL))
colnames(bray_dist_df)[1] <- "SampleID"
colnames(sample_data)[1]<-"SampleID"
bray_joined <- inner_join(bray_dist_df, sample_data, by = "SampleID") # combines the distance matrix with the metadata

bray_joined[1:5, 245:252]

```

```{r}

adonis2(formula = bray_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = bray_joined)
# not included tray effects yet as not all samples have a tray so it can mess with the model, will assess tray effects only on the sample types with trays/cages.

```

yes, there's overall significant clustering by sample type and building separately, and an interacting effect.

Next for pairwise.adonis to see where differences are. 

```{r}
pairwise.adonis2(bray_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```


#### now to subset the data to visualise and analyse between-insectary differences separately for each sample type


```{r}

#############  adults ############### 

ps_adult <- subset_samples(ps, Sample_Type2 == "AdultFemale")

adult_bray_dist = phyloseq::distance(ps_adult, method="bray", weighted=F)
adult_ordination = ordinate(ps_adult, method="PCoA", distance=adult_bray_dist) 

b_adult_plot <- plot_ordination(ps_adult, adult_ordination, color = "Building")  
b_adult_plot <- b_adult_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("adults")+
#  theme(legend.position="none")
  theme_bw()
  
b_adult_plot$layers <- b_adult_plot$layers[-1] 


############  larvae ############

ps_larvae <- subset_samples(ps, Sample_Type2 == "Larva")

larvae_bray_dist = phyloseq::distance(ps_larvae, method="bray", weighted=F)
larvae_ordination = ordinate(ps_larvae, method="PCoA", distance=larvae_bray_dist) 

b_larvae_plot <- plot_ordination(ps_larvae, larvae_ordination, color = "Building")  
b_larvae_plot <- b_larvae_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larvae")+
 # theme(legend.position="none")
  theme_bw()
b_larvae_plot$layers <- b_larvae_plot$layers[-1] 


###########  larval water ###########

ps_larvalwater <- subset_samples(ps, Sample_Type2 == "LarvalWater")

larvalwater_bray_dist = phyloseq::distance(ps_larvalwater, method="bray", weighted=F)
larvalwater_ordination = ordinate(ps_larvalwater, method="PCoA", distance=larvalwater_bray_dist) 

b_larvalwater_plot <- plot_ordination(ps_larvalwater, larvalwater_ordination, color = "Building")  
b_larvalwater_plot <- b_larvalwater_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larval water")+
#  theme(legend.position="none")
  theme_bw()
b_larvalwater_plot$layers <- b_larvalwater_plot$layers[-1]


###########  tap water ##############

ps_water <- subset_samples(ps, Sample_Type2 == "Water")

water_bray_dist = phyloseq::distance(ps_water, method="bray", weighted=F)
water_ordination = ordinate(ps_water, method="PCoA", distance=water_bray_dist) 

b_tap_water_plot <- plot_ordination(ps_water, water_ordination, color = "Building")  
b_tap_water_plot <- b_tap_water_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("tap water")+
  #theme(legend.position="none")
  theme_bw()
b_tap_water_plot$layers <- b_tap_water_plot$layers[-1]


########### fish food ############

ps_food <- subset_samples(ps, Sample_Type2 == "FishFood")

food_bray_dist = phyloseq::distance(ps_food, method="bray", weighted=F)
food_ordination = ordinate(ps_food, method="PCoA", distance=food_bray_dist) 

b_food_plot <- plot_ordination(ps_food, food_ordination, color = "Building")  
b_food_plot <- b_food_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("fish food")+
#  theme(legend.position="none")
  theme_bw()
b_food_plot$layers <- b_food_plot$layers[-1]


############ sugar ################

ps_sugar <- subset_samples(ps, Sample_Type2 == "Sugar")

sugar_bray_dist = phyloseq::distance(ps_sugar, method="bray", weighted=F)
sugar_ordination = ordinate(ps_sugar, method="PCoA", distance=sugar_bray_dist) 

b_sugar_plot <- plot_ordination(ps_sugar, sugar_ordination, color = "Building")  
b_sugar_plot <- b_sugar_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("sugar")+
  theme(legend.position="none") +
  theme_bw()
b_sugar_plot$layers <- b_sugar_plot$layers[-1]

```


show plots together

```{r}
pdf("bray_by_sampletype.pdf", width=7, height=5)
p2a <- ggpubr::ggarrange(b_tap_water_plot, b_food_plot , b_larvae_plot , b_larvalwater_plot , b_sugar_plot, b_adult_plot , ncol = 3, nrow = 2, common.legend = TRUE, legend="bottom")

p2a
dev.off()

```

```{r}
p2a
```



...an alternative version shows the original data plotted on the same axis scales, good for consistency, but don't see the same separation...


```{r}
alt_bray_facet_plots <- plot_ordination(ps, ordination, color = "Building") 
alt_bray_facet_plots <- alt_bray_facet_plots +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+ 
  facet_wrap(~Sample_Type2, nrow = 2, ncol = 3)
alt_bray_facet_plots$layers <- alt_bray_facet_plots$layers[-1] # this removed a first layer that was a solid point which had resulted in double overlaying points

alt_bray_facet_plots

```


Are there differences between buildings, within each sample type and are there tray/cage effects?


```{r}

#############  adults ############### 

adult_sample_data <- as.data.frame(sample_data(ps_adult))
adult_bray_dist_matrx = as.matrix(adult_bray_dist)
adult_bray_dist_df <- cbind(rownames(adult_bray_dist_matrx), data.frame(adult_bray_dist_matrx, row.names =NULL))
colnames(adult_bray_dist_df)[1] <- "SampleID"
colnames(adult_sample_data)[1]<-"SampleID"
adult_bray_joined <- inner_join(adult_bray_dist_df, adult_sample_data, by = "SampleID") # combines the distance matrix with the metadata

adonis2(formula = adult_bray_dist ~ Building + Building/Tray_Cage, data = adult_bray_joined) # this is looking at building effects, and also the nested trays within building effects

```
```{r}
pairwise.adonis2(adult_bray_dist ~ Building + Building/Tray_Cage, data = adult_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

#############  larvae ############### 

larvae_sample_data <- as.data.frame(sample_data(ps_larvae))
larvae_bray_dist_matrx = as.matrix(larvae_bray_dist)
larvae_bray_dist_df <- cbind(rownames(larvae_bray_dist_matrx), data.frame(larvae_bray_dist_matrx, row.names =NULL))
colnames(larvae_bray_dist_df)[1] <- "SampleID"
colnames(larvae_sample_data)[1]<-"SampleID"
larvae_bray_joined <- inner_join(larvae_bray_dist_df, larvae_sample_data, by = "SampleID") 

adonis2(formula = larvae_bray_dist ~ Building + Building/Tray_Cage, data = larvae_bray_joined) 

```
```{r}
pairwise.adonis2(larvae_bray_dist ~ Building + Building/Tray_Cage, data = larvae_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

#############  larval water ############### 

larvalwater_sample_data <- as.data.frame(sample_data(ps_larvalwater))
larvalwater_bray_dist_matrx = as.matrix(larvalwater_bray_dist)
larvalwater_bray_dist_df <- cbind(rownames(larvalwater_bray_dist_matrx), data.frame(larvalwater_bray_dist_matrx, row.names =NULL))
colnames(larvalwater_bray_dist_df)[1] <- "SampleID"
colnames(larvalwater_sample_data)[1]<-"SampleID"
larvalwater_bray_joined <- inner_join(larvalwater_bray_dist_df, larvalwater_sample_data, by = "SampleID") 

adonis2(formula = larvalwater_bray_dist ~ Building + Building/Tray_Cage, data = larvalwater_bray_joined) 

```
```{r}
pairwise.adonis2(larvalwater_bray_dist ~ Building + Building/Tray_Cage, data = larvalwater_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```



```{r}

#############  tap water ############### 

tapwater_sample_data <- as.data.frame(sample_data(ps_water))
tapwater_bray_dist_matrx = as.matrix(water_bray_dist)
tapwater_bray_dist_df <- cbind(rownames(tapwater_bray_dist_matrx), data.frame(tapwater_bray_dist_matrx, row.names =NULL))
colnames(tapwater_bray_dist_df)[1] <- "SampleID"
colnames(tapwater_sample_data)[1]<-"SampleID"
tapwater_bray_joined <- inner_join(tapwater_bray_dist_df, tapwater_sample_data, by = "SampleID") 

adonis2(formula = water_bray_dist ~ Building, data = tapwater_bray_joined) # no tray effects tested for with tap water

```

```{r}
pairwise.adonis2(water_bray_dist ~ Building, data = tapwater_bray_joined, p.adjust.method = "bonferroni", nperm = 999) 

```

```{r}

#############  fish food  ############### 

food_sample_data <- as.data.frame(sample_data(ps_food))
food_bray_dist_matrx = as.matrix(food_bray_dist)
food_bray_dist_df <- cbind(rownames(food_bray_dist_matrx), data.frame(food_bray_dist_matrx, row.names =NULL))
colnames(food_bray_dist_df)[1] <- "SampleID"
colnames(food_sample_data)[1]<-"SampleID"
food_bray_joined <- inner_join(food_bray_dist_df, food_sample_data, by = "SampleID") 

adonis2(formula = food_bray_dist ~ Building, data = food_bray_joined) # no tray effects tested for with food

```

```{r}
pairwise.adonis2(food_bray_dist ~ Building, data = food_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

#############  sugar solution ############### 

sugar_sample_data <- as.data.frame(sample_data(ps_sugar))
sugar_bray_dist_matrx = as.matrix(sugar_bray_dist)
sugar_bray_dist_df <- cbind(rownames(sugar_bray_dist_matrx), data.frame(sugar_bray_dist_matrx, row.names =NULL))
colnames(sugar_bray_dist_df)[1] <- "SampleID"
colnames(sugar_sample_data)[1]<-"SampleID"
sugar_bray_joined <- inner_join(sugar_bray_dist_df, sugar_sample_data, by = "SampleID") 

adonis2(formula = sugar_bray_dist ~ Building, data = sugar_bray_joined) 

```

```{r}
pairwise.adonis2(sugar_bray_dist ~ Building, data = sugar_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```



# unweighted unifrac

```{r}
uw_dist = phyloseq::distance(ps, method="unifrac", weighted=F) # calculate distances
uw_ordination = ordinate(ps, method="PCoA", distance=uw_dist) # ordinate

```

```{r}
# plot

uwordplot <- plot_ordination(ps, uw_ordination, color = "Sample_Type2") #basic plot

p3 <- uwordplot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Sample_Type2))+
  scale_fill_manual(values = colours)+
  scale_color_manual(values = colours)+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggtitle("unweighted Unifrac")


p3$layers <- p3$layers[-1] # this removed a first layer that was a solid point which had resulted in double overlaying points

p3  
```


plotting both uw and bray with all data


```{r}
pdf("bray_uw_by_sampletype.pdf", width=5, height=10)

p4 <- ggpubr::ggarrange(p2, p3, ncol = 1, nrow = 2, common.legend = TRUE, legend="bottom")
p4
dev.off()

```


overall differences by sample type and/or building. 

```{r}
uw_dist_matrx = as.matrix(uw_dist)
uw_dist_df <- cbind(rownames(uw_dist_matrx), data.frame(uw_dist_matrx, row.names =NULL))
colnames(uw_dist_df)[1] <- "SampleID"
uw_joined <- inner_join(uw_dist_df, sample_data, by = "SampleID") # combines the distance matrix with the metadata

adonis2(formula = uw_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = uw_joined)

```

```{r}
pairwise.adonis2(uw_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```



#### now to subset the data to visualise and analyse between-insectary differences separately for each sample type


```{r}

#############  adults ############### 

adult_uw_dist = phyloseq::distance(ps_adult, method="unifrac", weighted=F)
adult_uw_ordination = ordinate(ps_adult, method="PCoA", distance=adult_uw_dist) 

uw_adult_plot <- plot_ordination(ps_adult, adult_uw_ordination, color = "Building")  
uw_adult_plot <- uw_adult_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("adults")+
 # theme(legend.position="none")
  theme_bw()
uw_adult_plot$layers <- uw_adult_plot$layers[-1] 


############  larvae ############

larvae_uw_dist = phyloseq::distance(ps_larvae, method="unifrac", weighted=F)
larvae_uw_ordination = ordinate(ps_larvae, method="PCoA", distance=larvae_uw_dist) 

uw_larvae_plot <- plot_ordination(ps_larvae, larvae_uw_ordination, color = "Building")  
uw_larvae_plot <- uw_larvae_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larvae")+
 # theme(legend.position="none")
  theme_bw()
uw_larvae_plot$layers <- uw_larvae_plot$layers[-1] 


###########  larval water ###########

larvalwater_uw_dist = phyloseq::distance(ps_larvalwater, method="unifrac", weighted=F)
larvalwater_uw_ordination = ordinate(ps_larvalwater, method="PCoA", distance=larvalwater_uw_dist) 

uw_larvalwater_plot <- plot_ordination(ps_larvalwater, larvalwater_uw_ordination, color = "Building")  
uw_larvalwater_plot <- uw_larvalwater_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larval water")+
#  theme(legend.position="none")
  theme_bw()
uw_larvalwater_plot$layers <- uw_larvalwater_plot$layers[-1]


###########  tap water ##############

water_uw_dist = phyloseq::distance(ps_water, method="unifrac", weighted=F)
water_uw_ordination = ordinate(ps_water, method="PCoA", distance=water_uw_dist) 

uw_tap_water_plot <- plot_ordination(ps_water, water_uw_ordination, color = "Building")  
uw_tap_water_plot <- uw_tap_water_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("tap water")+
#  theme(legend.position="none")
  theme_bw()
uw_tap_water_plot$layers <- uw_tap_water_plot$layers[-1]


########### fish food ############

food_uw_dist = phyloseq::distance(ps_food, method="unifrac", weighted=F)
food_uw_ordination = ordinate(ps_food, method="PCoA", distance=food_uw_dist) 

uw_food_plot <- plot_ordination(ps_food, food_uw_ordination, color = "Building")  
uw_food_plot <- uw_food_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("fish food")+
 # theme(legend.position="none")
  theme_bw()
uw_food_plot$layers <- uw_food_plot$layers[-1]


############ sugar ################

sugar_uw_dist = phyloseq::distance(ps_sugar, method="unifrac", weighted=F)
sugar_uw_ordination = ordinate(ps_sugar, method="PCoA", distance=sugar_uw_dist) 

uw_sugar_plot <- plot_ordination(ps_sugar, sugar_uw_ordination, color = "Building")  
uw_sugar_plot <- uw_sugar_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("sugar")+
  #theme(legend.position="none")+
  theme_bw()
uw_sugar_plot$layers <- uw_sugar_plot$layers[-1]

```


show plots together
 

```{r}
pdf("uw_by_sampletype.pdf", width=7, height=5)
p3a <- ggpubr::ggarrange(uw_tap_water_plot, uw_food_plot , uw_larvae_plot , uw_larvalwater_plot , uw_sugar_plot, uw_adult_plot , ncol = 3, nrow = 2, common.legend = TRUE, legend="bottom")

p3a
dev.off()

```

```{r}
p3a
```

Are there differences between buildings, within each sample type and are there tray/cage effects?


```{r}

#############  adults ############### 

adult_uw_dist_matrx = as.matrix(adult_uw_dist)
adult_uw_dist_df <- cbind(rownames(adult_uw_dist_matrx), data.frame(adult_uw_dist_matrx, row.names =NULL))
colnames(adult_uw_dist_df)[1] <- "SampleID"
adult_uw_joined <- inner_join(adult_uw_dist_df, adult_sample_data, by = "SampleID") 

adonis2(formula = adult_uw_dist ~ Building + Building/Tray_Cage, data = adult_uw_joined) 
```

```{r}
pairwise.adonis2(adult_uw_dist ~ Building + Building/Tray_Cage, data = adult_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```


```{r}

#############  larvae ############### 

larvae_uw_dist_matrx = as.matrix(larvae_uw_dist)
larvae_uw_dist_df <- cbind(rownames(larvae_uw_dist_matrx), data.frame(larvae_uw_dist_matrx, row.names =NULL))
colnames(larvae_uw_dist_df)[1] <- "SampleID"
larvae_uw_joined <- inner_join(larvae_uw_dist_df, larvae_sample_data, by = "SampleID") 

adonis2(formula = larvae_uw_dist ~ Building + Building/Tray_Cage, data = larvae_uw_joined) 

```

```{r}
pairwise.adonis2(larvae_uw_dist ~ Building + Building/Tray_Cage, data = larvae_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```


```{r}

#############  larval water ############### 

larvalwater_uw_dist_matrx = as.matrix(larvalwater_uw_dist)
larvalwater_uw_dist_df <- cbind(rownames(larvalwater_uw_dist_matrx), data.frame(larvalwater_uw_dist_matrx, row.names =NULL))
colnames(larvalwater_uw_dist_df)[1] <- "SampleID"
larvalwater_uw_joined <- inner_join(larvalwater_uw_dist_df, larvalwater_sample_data, by = "SampleID") 

adonis2(formula = larvalwater_uw_dist ~ Building + Building/Tray_Cage, data = larvalwater_uw_joined) 

```

```{r}
pairwise.adonis2(larvalwater_uw_dist ~ Building + Building/Tray_Cage, data = larvalwater_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

#############  tap water ############### 

tapwater_uw_dist_matrx = as.matrix(water_uw_dist)
tapwater_uw_dist_df <- cbind(rownames(tapwater_uw_dist_matrx), data.frame(tapwater_uw_dist_matrx, row.names =NULL))
colnames(tapwater_uw_dist_df)[1] <- "SampleID"
tapwater_uw_joined <- inner_join(tapwater_uw_dist_df, tapwater_sample_data, by = "SampleID") 

adonis2(formula = water_uw_dist ~ Building, data = tapwater_uw_joined) # no tray effects tested for with tap water

```

```{r}
pairwise.adonis2(water_uw_dist ~ Building, data = tapwater_uw_joined, p.adjust.method = "bonferroni", nperm = 999)


```


```{r}

#############  fish food  ############### 

food_uw_dist_matrx = as.matrix(food_uw_dist)
food_uw_dist_df <- cbind(rownames(food_uw_dist_matrx), data.frame(food_uw_dist_matrx, row.names =NULL))
colnames(food_uw_dist_df)[1] <- "SampleID"
food_uw_joined <- inner_join(food_uw_dist_df, food_sample_data, by = "SampleID") 

adonis2(formula = food_uw_dist ~ Building, data = food_uw_joined) # no tray effects tested for with food

```
not statistically significant

```{r}

#############  sugar solution ############### 

sugar_uw_dist_matrx = as.matrix(sugar_uw_dist)
sugar_uw_dist_df <- cbind(rownames(sugar_uw_dist_matrx), data.frame(sugar_uw_dist_matrx, row.names =NULL))
colnames(sugar_uw_dist_df)[1] <- "SampleID"
sugar_uw_joined <- inner_join(sugar_uw_dist_df, sugar_sample_data, by = "SampleID") 

adonis2(formula = sugar_uw_dist ~ Building, data = sugar_uw_joined) 

```

```{r}
pairwise.adonis2(food_uw_dist ~ Building, data = food_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```


