---
title: "Ins_comp_7-survival"
author: "laura"
date: "2024-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.


https://bioconductor.org/packages/release/bioc/vignettes/MicrobiotaProcess/inst/doc//MicrobiotaProcess.html

can also look at https://www.yanh.org/2021/01/01/microbiome-r/

# Install packages

```{r}
library(ggplot2)
```

# import data

```{r}
dat = read.csv("ins_comp_survival_data.csv")


sapply(dat, class)

as.numeric(dat$number)


```

```{r}
colours2 <- c(
  "ACC" = "#00A087FF",
  "CTID" = "#3C5488FF",
  "MK" = "#9C58A1"
)
```

# plot 

```{r}

p <- ggplot(data= dat, aes(x= tray, y= number, colour = insectary, shape = stage)) +
  geom_point(size = 3) + 
   scale_color_manual(values = colours2) +
  scale_y_continuous(limits = c(90, 150)) +
  facet_grid(~factor(insectary, levels=c("CTID", "MK", "ACC")), scales = "free", space = "free")
p


```

# stats

using the same data presented in a slightly different way


```{r}

# import data
dat2 = read.csv("ins_comp_survival2.csv")

# sapply(dat2, class)

```

```{r}

# ANOVA
anova_pupae <- aov(stage2 ~ insectary, data = dat2)
anova_adult <- aov(stage3 ~ insectary, data = dat2)

summary(anova_pupae)
summary(anova_adult)


# Kruskal-WAllis
Kruskal_pupae <- kruskal.test(stage2 ~ insectary, data = dat2)
Kruskal_adult <- kruskal.test(stage3 ~ insectary, data = dat2)


if (Kruskal_pupae$p.value < 0.05) {
  print("There are significant differences between at least two groups.")
} else {
  print("There is not enough evidence to conclude significant differences between groups.")
}

if (Kruskal_adult$p.value < 0.05) {
  print("There are significant differences between at least two groups.")
} else {
  print("There is not enough evidence to conclude significant differences between groups.")
}


```

No statistically significant difference in numbers making it to pupal and adult stages in the different rooms, but only 3 reps per room so it would need to be a big difference to see that.


# pupal development time


```{r}
pup = read.csv("pupae_development.csv")

```

# plot 

```{r}

p2 <- ggplot(data= pup, aes(x= time.to.pupation, y= number, colour = insectary, shape = tray)) +
  geom_point(size = 3) + 
   scale_color_manual(values = colours2) +
  scale_y_continuous(limits = c(0, 150)) +
  facet_grid(~factor(insectary, levels=c("CTID", "MK", "ACC")), scales = "free", space = "free")+
  geom_line()
p2


```

