---
title: "Ins_comp_5-ASVs"
author: "laura"
date: "2024-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.


Now I will try a couple of ways to investigate whether there are signatures or particular ASVs associated with different insectaries and in which sample types these appear - is it larvae and larval water as suggested from diversity plot groupings or is it others, or do signature ASVs persist?

# Install packages

```{r}
# phyloseq itself
#if(!requireNamespace("BiocManager")){
#  install.packages("BiocManager")
#   }
#BiocManager::install("phyloseq")


library(phyloseq)
library(ggplot2)
library(dplyr) 
library(pheatmap)
```

# import data

```{r}

rm(list = ls()) 
ps = readRDS("physeq_obj_3974") 
head(sample_data(ps))

```


#  following phyloseq demo

https://joey711.github.io/phyloseq-demo/phyloseq-demo-slides.html
also useful : https://microucph.github.io/amplicon_data_analysis/html/phyloseq_operations.html
https://github.com/joey711/phyloseq/issues/1089



```{r}

# transform counts to rel abund
ps2 <- transform_sample_counts(ps, function(x) x/sum(x))

#Remove taxa not seen more than 0.05 rel abund in at least 0.004% of the samples (this equals 1 sample as there are 244 in total). This protects against an OTU with small mean & trivially large C.V. (I can't work out how to say - in any one sample)

ps3 = filter_taxa(ps2, function(x) sum(x >= 0.05) >= (0.004*length(x)), TRUE) # this has 34 ASVs



```


### converting the zeros to NAs for convenient plotting


```{r}
# take components out of phloseq obj

tax <- phyloseq::tax_table(ps3)
meta <- sample_data(ps3)
asv <- otu_table(ps3)

asvdf <- as.data.frame(asv)
asvdf[asvdf == 0] <- NA 
  

asv2 <- as.matrix(asvdf) # convert to matrix
asv2 = otu_table(asv2, taxa_are_rows = TRUE) # convert to necessary format  to make phyloseq object


psNAs <- phyloseq(asv2, tax, meta)  

```


To remember what families the ASVs belong to...

```{r}
heatmap_test10a <- plot_heatmap(
              psNAs, 
              method = NULL,
              taxa.label = "Genus",
             taxa.order = "Family",
             sample.order = "Building",
             sample.label = "Tray_Cage",
             trans = NULL,
            low = "#B1C9E8",
            high = "#002F6C",
             na.value = "white",
             title = "relative abundance of all ASVs seen at >= 5% reads in >= 1 sample")

heatmap_test10a


```


```{r}

# to see with tray/cage IDs for plotting

heatmap_test11a <- heatmap_test10a + facet_grid(~factor(Sample_Type2, levels=c("Water", "FishFood", "LarvalWater", "Larva", "Sugar", "AdultFemale")), scales = "free_x", space = "free_x") 

heatmap_test11a 


#pdf("ASV_heatmap_tray_cage.pdf", width=12, height=6)
heatmap_test11a 
#dev.off()


```

