---
title: "Ins_comp_3c-beta_diversity_metrics"
author: "laura"
date: "2024-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.

Analysis of beta diversity - bray curtis and unweighted unifrac


### Install packages

```{r}
 
library(phyloseq)
library(ggplot2)
library(dplyr) 
library(ggpubr)
library(dunn.test)
library(pairwiseAdonis)
library(vegan)

```

### import data

```{r}
ps = readRDS("physeq_obj_3974") 

```

### the colours I'll use

```{r}
colours <- c(
  "AdultFemale" = "#56187D", 
  "FishFood" = "#AC0123", 
  "Larva" = "#006347", 
  "LarvalWater" = "#97D0A7", 
  "Sugar" = "#EF9F26", 
  "Sugar2" = "#E6CD69", 
  "Water" = "#7AA8D7"
)


colours2 <- c(
  "C-ACC" = "#00A087FF",
  "A-CTID" = "#3C5488FF",
  "B-MK" = "#9C58A1"
)

```

## beta diversity

Not doing Hellinger transformation (as is suggested by the microbiotaProcess tutorial/vignette) as the input data are rarefied

# bray curtis first

### firstly, all data together


```{r}
bray_dist = phyloseq::distance(ps, method="bray", weighted=F) # calculate distances
ordination = ordinate(ps, method="PCoA", distance=bray_dist) # ordinate

```

```{r}
# plot

brayordplot <- plot_ordination(ps, ordination, color = "Sample_Type2") #basic plot

p2 <- brayordplot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Sample_Type2))+
  scale_fill_manual(values = colours)+
  scale_color_manual(values = colours)+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggtitle("bray-curtis")


p2$layers <- p2$layers[-1] 

p2  
```
overall differences by sample type and/or building. 

First, easier if we take the data out of the phyloseq object


```{r}
sample_data <- as.data.frame(sample_data(ps))
bray_dist_matrx = as.matrix(bray_dist)
bray_dist_df <- cbind(rownames(bray_dist_matrx), data.frame(bray_dist_matrx, row.names =NULL))
colnames(bray_dist_df)[1] <- "SampleID"
colnames(sample_data)[1]<-"SampleID"
bray_joined <- inner_join(bray_dist_df, sample_data, by = "SampleID") # combines the distance matrix with the metadata

bray_joined[1:5, 245:252]

```

```{r}

adonis2(formula = bray_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = bray_joined)
# not included tray effects yet as not all samples have a tray so it can mess with the model, will assess tray effects only on the sample types with trays/cages.

```

yes, there's overall significant clustering by sample type and building separately, and an interacting effect.

Next for pairwise.adonis to see where differences are. 

```{r}
pairwise.adonis2(bray_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```


next, use PERMDISP (permutational analysis of multivariate dispersions) to test whether the differences are due to differences in dispersion (variance) or locations of controids 


```{r}

# Run betadisper for each grouping variable
# For Sample_Type2
disp_sample_type <- betadisper(bray_dist, bray_joined$Sample_Type2)

# For Building
disp_building <- betadisper(bray_dist, bray_joined$Building)

# For the interaction Sample_Type2:Building
disp_interaction <- betadisper(bray_dist, interaction(bray_joined$Sample_Type2, bray_joined$Building))


```


```{r}
# Test for dispersion differences in Sample_Type2
perm_disp_sample_type <- permutest(disp_sample_type, permutations = 999)
print(perm_disp_sample_type)
```

significant (p < 0.05), it suggests that there may be differences in the spread (dispersion) of samples within groups, which could potentially explain some of the differences detected by adonis2. If results were not significant (p > 0.05), it would suggest that differences found with adonis2 are more likely due to differences in the location of centroids rather than dispersion.

```{r}
# Test for dispersion differences in Building
perm_disp_building <- permutest(disp_building, permutations = 999)
print(perm_disp_building)
```


```{r}
# Test for dispersion differences in the interaction Sample_Type2:Building
perm_disp_interaction <- permutest(disp_interaction, permutations = 999)
print(perm_disp_interaction)

```
significant


#### now to subset the data to visualise and analyse between-insectary differences separately for each sample type


```{r}

#############  adults ############### 

ps_adult <- subset_samples(ps, Sample_Type2 == "AdultFemale")

adult_bray_dist = phyloseq::distance(ps_adult, method="bray", weighted=F)
adult_ordination = ordinate(ps_adult, method="PCoA", distance=adult_bray_dist) 

b_adult_plot <- plot_ordination(ps_adult, adult_ordination, color = "Building")  
b_adult_plot <- b_adult_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("adults")+
  theme_bw()
  
b_adult_plot$layers <- b_adult_plot$layers[-1] 


############  larvae ############

ps_larvae <- subset_samples(ps, Sample_Type2 == "Larva")

larvae_bray_dist = phyloseq::distance(ps_larvae, method="bray", weighted=F)
larvae_ordination = ordinate(ps_larvae, method="PCoA", distance=larvae_bray_dist) 

b_larvae_plot <- plot_ordination(ps_larvae, larvae_ordination, color = "Building")  
b_larvae_plot <- b_larvae_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larvae")+
  theme_bw()
b_larvae_plot$layers <- b_larvae_plot$layers[-1] 


###########  larval water ###########

ps_larvalwater <- subset_samples(ps, Sample_Type2 == "LarvalWater")

larvalwater_bray_dist = phyloseq::distance(ps_larvalwater, method="bray", weighted=F)
larvalwater_ordination = ordinate(ps_larvalwater, method="PCoA", distance=larvalwater_bray_dist) 

b_larvalwater_plot <- plot_ordination(ps_larvalwater, larvalwater_ordination, color = "Building")  
b_larvalwater_plot <- b_larvalwater_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larval water")+
  theme_bw()
b_larvalwater_plot$layers <- b_larvalwater_plot$layers[-1]


###########  tap water ##############

ps_water <- subset_samples(ps, Sample_Type2 == "Water")

water_bray_dist = phyloseq::distance(ps_water, method="bray", weighted=F)
water_ordination = ordinate(ps_water, method="PCoA", distance=water_bray_dist) 

b_tap_water_plot <- plot_ordination(ps_water, water_ordination, color = "Building")  
b_tap_water_plot <- b_tap_water_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("tap water")+
  theme_bw()
b_tap_water_plot$layers <- b_tap_water_plot$layers[-1]


########### fish food ############

ps_food <- subset_samples(ps, Sample_Type2 == "FishFood")

food_bray_dist = phyloseq::distance(ps_food, method="bray", weighted=F)
food_ordination = ordinate(ps_food, method="PCoA", distance=food_bray_dist) 

b_food_plot <- plot_ordination(ps_food, food_ordination, color = "Building")  
b_food_plot <- b_food_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("fish food")+
  theme_bw()
b_food_plot$layers <- b_food_plot$layers[-1]


############ sugar ################

ps_sugar <- subset_samples(ps, Sample_Type2 == "Sugar")

sugar_bray_dist = phyloseq::distance(ps_sugar, method="bray", weighted=F)
sugar_ordination = ordinate(ps_sugar, method="PCoA", distance=sugar_bray_dist) 

b_sugar_plot <- plot_ordination(ps_sugar, sugar_ordination, color = "Building")  
b_sugar_plot <- b_sugar_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("sugar")+
  theme(legend.position="none") +
  theme_bw()
b_sugar_plot$layers <- b_sugar_plot$layers[-1]

```


show plots together

```{r}
p2a <- ggpubr::ggarrange(b_tap_water_plot, b_food_plot , b_larvae_plot , b_larvalwater_plot , b_sugar_plot, b_adult_plot , ncol = 3, nrow = 2, common.legend = TRUE, legend="bottom")

p2a

```



Are there differences between buildings, within each sample type and are there tray/cage effects?


```{r}

#############  adults ############### 

adult_sample_data <- as.data.frame(sample_data(ps_adult))
adult_bray_dist_matrx = as.matrix(adult_bray_dist)
adult_bray_dist_df <- cbind(rownames(adult_bray_dist_matrx), data.frame(adult_bray_dist_matrx, row.names =NULL))
colnames(adult_bray_dist_df)[1] <- "SampleID"
colnames(adult_sample_data)[1]<-"SampleID"
adult_bray_joined <- inner_join(adult_bray_dist_df, adult_sample_data, by = "SampleID") 

adonis2(formula = adult_bray_dist ~ Building + Building/Tray_Cage, data = adult_bray_joined) # this is looking at building effects, and also the nested trays within building effects

```

```{r}
pairwise.adonis2(adult_bray_dist ~ Building + Building/Tray_Cage, data = adult_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```


```{r}

# Run betadisper for Building
adult_disp_building <- betadisper(adult_bray_dist, adult_bray_joined$Building)

# Test for dispersion differences in Building
adult_perm_disp_building <- permutest(adult_disp_building, permutations = 999)
print(adult_perm_disp_building)

```

significant



```{r}

# Define the buildings and corresponding labels
buildings <- c("A-CTID", "B-MK", "C-ACC")
labels <- c("adult_A", "adult_B", "adult_C")

# Initialize lists to store results
bray_dists <- list()
sample_data_frames <- list()
dist_matrices <- list()
joined_dfs <- list()
disp_Tray_Cage <- list()
perm_disp_Tray_Cage <- list()

# Loop over each building
for (i in seq_along(buildings)) {
  # Subset samples for the current building
  current_building <- buildings[i]
  current_label <- labels[i]
  
  # Dynamically assign variable names
  subset_data <- subset_samples(ps_adult, Building == current_building)
  
  # Calculate the Bray-Curtis distance
  bray_dist <- phyloseq::distance(subset_data, method="bray", weighted=FALSE)
  
  # Get sample data and convert to a data frame
  sample_data_df <- as.data.frame(sample_data(subset_data))
  
  # Convert Bray-Curtis distance to matrix and data frame
  bray_dist_matrix <- as.matrix(bray_dist)
  bray_dist_df <- cbind(rownames(bray_dist_matrix), data.frame(bray_dist_matrix, row.names=NULL))
  colnames(bray_dist_df)[1] <- "SampleID"
  colnames(sample_data_df)[1] <- "SampleID"
  
  # Join the distance data frame with sample data
  bray_joined <- inner_join(bray_dist_df, sample_data_df, by="SampleID")
  
  # Beta dispersion and permutation test
  disp_Tray_Cage_result <- betadisper(bray_dist, bray_joined$Tray_Cage)
  perm_disp_Tray_Cage_result <- permutest(disp_Tray_Cage_result, permutations=999)
  
  # Print the results for each building
  cat("Results for Building:", current_building, "\n")
  print(perm_disp_Tray_Cage_result)
  
  # Store results in lists
  bray_dists[[current_label]] <- bray_dist
  sample_data_frames[[current_label]] <- sample_data_df
  dist_matrices[[current_label]] <- bray_dist_matrix
  joined_dfs[[current_label]] <- bray_joined
  disp_Tray_Cage[[current_label]] <- disp_Tray_Cage_result
  perm_disp_Tray_Cage[[current_label]] <- perm_disp_Tray_Cage_result
}

```

building significant for adults (both adonis2 and permdisp. not tray/cage or interaction (although at tray/cage level numbers of reps are getting small), tray/cage results not significant with permdisp either.

```{r}

#############  larvae ############### 

larvae_sample_data <- as.data.frame(sample_data(ps_larvae))
larvae_bray_dist_matrx = as.matrix(larvae_bray_dist)
larvae_bray_dist_df <- cbind(rownames(larvae_bray_dist_matrx), data.frame(larvae_bray_dist_matrx, row.names =NULL))
colnames(larvae_bray_dist_df)[1] <- "SampleID"
colnames(larvae_sample_data)[1]<-"SampleID"
larvae_bray_joined <- inner_join(larvae_bray_dist_df, larvae_sample_data, by = "SampleID") 

adonis2(formula = larvae_bray_dist ~ Building + Building/Tray_Cage, data = larvae_bray_joined) 

```

```{r}
pairwise.adonis2(larvae_bray_dist ~ Building + Building/Tray_Cage, data = larvae_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

larvae_disp_building <- betadisper(larvae_bray_dist, larvae_bray_joined$Building)

larvae_perm_disp_building <- permutest(larvae_disp_building, permutations = 999)
print(larvae_perm_disp_building) # signif

```

```{r}

L_bray_dists <- list()
L_sample_data_frames <- list()
L_dist_matrices <- list()
L_joined_dfs <- list()
L_disp_Tray_Cage <- list()
L_perm_disp_Tray_Cage <- list()

for (i in seq_along(buildings)) {
    current_building <- buildings[i]
  current_label <- labels[i]
  
  L_subset_data <- subset_samples(ps_larvae, Building == current_building)
  
  L_bray_dist <- phyloseq::distance(L_subset_data, method="bray", weighted=FALSE)
  
  L_sample_data_df <- as.data.frame(sample_data(L_subset_data))
  
  L_bray_dist_matrix <- as.matrix(L_bray_dist)
  L_bray_dist_df <- cbind(rownames(L_bray_dist_matrix), data.frame(L_bray_dist_matrix, row.names=NULL))
  colnames(L_bray_dist_df)[1] <- "SampleID"
  colnames(L_sample_data_df)[1] <- "SampleID"
  
  L_bray_joined <- inner_join(L_bray_dist_df, L_sample_data_df, by="SampleID")
  
  L_disp_Tray_Cage_result <- betadisper(L_bray_dist, L_bray_joined$Tray_Cage)
  L_perm_disp_Tray_Cage_result <- permutest(L_disp_Tray_Cage_result, permutations=999)
  
  cat("Results for Building Larvae:", current_building, "\n")
  print(L_perm_disp_Tray_Cage_result)
  
  L_bray_dists[[current_label]] <- L_bray_dist
  L_sample_data_frames[[current_label]] <- L_sample_data_df
  L_dist_matrices[[current_label]] <- L_bray_dist_matrix
  L_joined_dfs[[current_label]] <- L_bray_joined
  L_disp_Tray_Cage[[current_label]] <- L_disp_Tray_Cage_result
  L_perm_disp_Tray_Cage[[current_label]] <- L_perm_disp_Tray_Cage_result
}

```
none significant



```{r}

#############  larval water ############### 

larvalwater_sample_data <- as.data.frame(sample_data(ps_larvalwater))
larvalwater_bray_dist_matrx = as.matrix(larvalwater_bray_dist)
larvalwater_bray_dist_df <- cbind(rownames(larvalwater_bray_dist_matrx), data.frame(larvalwater_bray_dist_matrx, row.names =NULL))
colnames(larvalwater_bray_dist_df)[1] <- "SampleID"
colnames(larvalwater_sample_data)[1]<-"SampleID"
larvalwater_bray_joined <- inner_join(larvalwater_bray_dist_df, larvalwater_sample_data, by = "SampleID") 

adonis2(formula = larvalwater_bray_dist ~ Building + Building/Tray_Cage, data = larvalwater_bray_joined) 

```

```{r}
pairwise.adonis2(larvalwater_bray_dist ~ Building + Building/Tray_Cage, data = larvalwater_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

larvalwater_disp_building <- betadisper(larvalwater_bray_dist, larvalwater_bray_joined$Building)
larvalwater_perm_disp_building <- permutest(larvalwater_disp_building, permutations = 999)
print(larvalwater_perm_disp_building)

```


```{r}

LW_bray_dists <- list()
LW_sample_data_frames <- list()
LW_dist_matrices <- list()
LW_joined_dfs <- list()
LW_disp_Tray_Cage <- list()
LW_perm_disp_Tray_Cage <- list()

for (i in seq_along(buildings)) {
  current_building <- buildings[i]
  current_label <- labels[i]
  
  LW_subset_data <- subset_samples(ps_larvalwater, Building == current_building)
  
  LW_bray_dist <- phyloseq::distance(LW_subset_data, method="bray", weighted=FALSE)
  
  LW_sample_data_df <- as.data.frame(sample_data(LW_subset_data))
  
  LW_bray_dist_matrix <- as.matrix(LW_bray_dist)
  LW_bray_dist_df <- cbind(rownames(LW_bray_dist_matrix), data.frame(LW_bray_dist_matrix, row.names=NULL))
  colnames(LW_bray_dist_df)[1] <- "SampleID"
  colnames(LW_sample_data_df)[1] <- "SampleID"
  
  LW_bray_joined <- inner_join(LW_bray_dist_df, LW_sample_data_df, by="SampleID")
  
  LW_disp_Tray_Cage_result <- betadisper(LW_bray_dist, LW_bray_joined$Tray_Cage)
  LW_perm_disp_Tray_Cage_result <- permutest(LW_disp_Tray_Cage_result, permutations=999)
  
  cat("Results for Building Larval water:", current_building, "\n")
  print(LW_perm_disp_Tray_Cage_result)
  
  LW_bray_dists[[current_label]] <- LW_bray_dist
  LW_sample_data_frames[[current_label]] <- LW_sample_data_df
  LW_dist_matrices[[current_label]] <- LW_bray_dist_matrix
  LW_joined_dfs[[current_label]] <- LW_bray_joined
  LW_disp_Tray_Cage[[current_label]] <- LW_disp_Tray_Cage_result
  LW_perm_disp_Tray_Cage[[current_label]] <- LW_perm_disp_Tray_Cage_result
}

```
non significant

```{r}

#############  tap water ############### 

tapwater_sample_data <- as.data.frame(sample_data(ps_water))
tapwater_bray_dist_matrx = as.matrix(water_bray_dist)
tapwater_bray_dist_df <- cbind(rownames(tapwater_bray_dist_matrx), data.frame(tapwater_bray_dist_matrx, row.names =NULL))
colnames(tapwater_bray_dist_df)[1] <- "SampleID"
colnames(tapwater_sample_data)[1]<-"SampleID"
tapwater_bray_joined <- inner_join(tapwater_bray_dist_df, tapwater_sample_data, by = "SampleID") 

adonis2(formula = water_bray_dist ~ Building, data = tapwater_bray_joined) # no tray effects tested for with tap water

```

```{r}
pairwise.adonis2(water_bray_dist ~ Building, data = tapwater_bray_joined, p.adjust.method = "bonferroni", nperm = 999) 

```

```{r}

tapwater_disp_building <- betadisper(water_bray_dist, tapwater_bray_joined$Building)
tapwater_perm_disp_building <- permutest(tapwater_disp_building, permutations = 999)
print(tapwater_perm_disp_building)

```

```{r}

#############  fish food  ############### 

food_sample_data <- as.data.frame(sample_data(ps_food))
food_bray_dist_matrx = as.matrix(food_bray_dist)
food_bray_dist_df <- cbind(rownames(food_bray_dist_matrx), data.frame(food_bray_dist_matrx, row.names =NULL))
colnames(food_bray_dist_df)[1] <- "SampleID"
colnames(food_sample_data)[1]<-"SampleID"
food_bray_joined <- inner_join(food_bray_dist_df, food_sample_data, by = "SampleID") 

adonis2(formula = food_bray_dist ~ Building, data = food_bray_joined) # no tray effects tested for with food

```

```{r}
pairwise.adonis2(food_bray_dist ~ Building, data = food_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

food_disp_building <- betadisper(food_bray_dist, food_bray_joined$Building)
food_perm_disp_building <- permutest(food_disp_building, permutations = 999)
print(food_perm_disp_building)

```

```{r}

#############  sugar solution ############### 

sugar_sample_data <- as.data.frame(sample_data(ps_sugar))
sugar_bray_dist_matrx = as.matrix(sugar_bray_dist)
sugar_bray_dist_df <- cbind(rownames(sugar_bray_dist_matrx), data.frame(sugar_bray_dist_matrx, row.names =NULL))
colnames(sugar_bray_dist_df)[1] <- "SampleID"
colnames(sugar_sample_data)[1]<-"SampleID"
sugar_bray_joined <- inner_join(sugar_bray_dist_df, sugar_sample_data, by = "SampleID") 

adonis2(formula = sugar_bray_dist ~ Building, data = sugar_bray_joined) 

```

```{r}
pairwise.adonis2(sugar_bray_dist ~ Building, data = sugar_bray_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

sugar_disp_building <- betadisper(sugar_bray_dist, sugar_bray_joined$Building)
sugar_perm_disp_building <- permutest(sugar_disp_building, permutations = 999)
print(sugar_perm_disp_building)

```



# unweighted unifrac

```{r}
uw_dist = phyloseq::distance(ps, method="unifrac", weighted=F) 
uw_ordination = ordinate(ps, method="PCoA", distance=uw_dist) 

```

```{r}

uwordplot <- plot_ordination(ps, uw_ordination, color = "Sample_Type2") 

p3 <- uwordplot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Sample_Type2))+
  scale_fill_manual(values = colours)+
  scale_color_manual(values = colours)+
  theme_bw()+
  theme(legend.position = "bottom")+
  ggtitle("unweighted Unifrac")


p3$layers <- p3$layers[-1] 

p3  
```


overall differences by sample type and/or building. 

```{r}
uw_dist_matrx = as.matrix(uw_dist)
uw_dist_df <- cbind(rownames(uw_dist_matrx), data.frame(uw_dist_matrx, row.names =NULL))
colnames(uw_dist_df)[1] <- "SampleID"
uw_joined <- inner_join(uw_dist_df, sample_data, by = "SampleID") 

adonis2(formula = uw_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = uw_joined)

```

```{r}
pairwise.adonis2(uw_dist ~ Sample_Type2 + Building + Sample_Type2:Building, data = uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

uw_disp_sample_type <- betadisper(uw_dist, uw_joined$Sample_Type2)
uw_disp_building <- betadisper(uw_dist, uw_joined$Building)
uw_disp_interaction <- betadisper(uw_dist, interaction(uw_joined$Sample_Type2, uw_joined$Building))

uw_perm_disp_sample_type <- permutest(uw_disp_sample_type, permutations = 999)
print(uw_perm_disp_sample_type) # signif 
uw_perm_disp_building <- permutest(uw_disp_building, permutations = 999)
print(uw_perm_disp_building) # not signif
uw_perm_disp_interaction <- permutest(uw_disp_interaction, permutations = 999)
print(uw_perm_disp_interaction) # signif
```

#### now to subset the data to visualise and analyse between-insectary differences separately for each sample type


```{r}

#############  adults ############### 

adult_uw_dist = phyloseq::distance(ps_adult, method="unifrac", weighted=F)
adult_uw_ordination = ordinate(ps_adult, method="PCoA", distance=adult_uw_dist) 

uw_adult_plot <- plot_ordination(ps_adult, adult_uw_ordination, color = "Building")  
uw_adult_plot <- uw_adult_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("adults")+
  theme_bw()
uw_adult_plot$layers <- uw_adult_plot$layers[-1] 


############  larvae ############

larvae_uw_dist = phyloseq::distance(ps_larvae, method="unifrac", weighted=F)
larvae_uw_ordination = ordinate(ps_larvae, method="PCoA", distance=larvae_uw_dist) 

uw_larvae_plot <- plot_ordination(ps_larvae, larvae_uw_ordination, color = "Building")  
uw_larvae_plot <- uw_larvae_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larvae")+
  theme_bw()
uw_larvae_plot$layers <- uw_larvae_plot$layers[-1] 


###########  larval water ###########

larvalwater_uw_dist = phyloseq::distance(ps_larvalwater, method="unifrac", weighted=F)
larvalwater_uw_ordination = ordinate(ps_larvalwater, method="PCoA", distance=larvalwater_uw_dist) 

uw_larvalwater_plot <- plot_ordination(ps_larvalwater, larvalwater_uw_ordination, color = "Building")  
uw_larvalwater_plot <- uw_larvalwater_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("larval water")+
  theme_bw()
uw_larvalwater_plot$layers <- uw_larvalwater_plot$layers[-1]


###########  tap water ##############

water_uw_dist = phyloseq::distance(ps_water, method="unifrac", weighted=F)
water_uw_ordination = ordinate(ps_water, method="PCoA", distance=water_uw_dist) 

uw_tap_water_plot <- plot_ordination(ps_water, water_uw_ordination, color = "Building")  
uw_tap_water_plot <- uw_tap_water_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("tap water")+
  theme_bw()
uw_tap_water_plot$layers <- uw_tap_water_plot$layers[-1]


########### fish food ############

food_uw_dist = phyloseq::distance(ps_food, method="unifrac", weighted=F)
food_uw_ordination = ordinate(ps_food, method="PCoA", distance=food_uw_dist) 

uw_food_plot <- plot_ordination(ps_food, food_uw_ordination, color = "Building")  
uw_food_plot <- uw_food_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("fish food")+
  theme_bw()
uw_food_plot$layers <- uw_food_plot$layers[-1]


############ sugar ################

sugar_uw_dist = phyloseq::distance(ps_sugar, method="unifrac", weighted=F)
sugar_uw_ordination = ordinate(ps_sugar, method="PCoA", distance=sugar_uw_dist) 

uw_sugar_plot <- plot_ordination(ps_sugar, sugar_uw_ordination, color = "Building")  
uw_sugar_plot <- uw_sugar_plot +
  geom_point(size=4, alpha=0.5)+
  stat_ellipse(geom = "polygon", alpha = 0.25, aes(fill = Building))+
  scale_fill_manual(values = colours2)+
  scale_color_manual(values = colours2)+
  ggtitle("sugar")+
   theme_bw()
uw_sugar_plot$layers <- uw_sugar_plot$layers[-1]

```


show plots together
 

```{r}
p3a <- ggpubr::ggarrange(uw_tap_water_plot, uw_food_plot , uw_larvae_plot , uw_larvalwater_plot , uw_sugar_plot, uw_adult_plot , ncol = 3, nrow = 2, common.legend = TRUE, legend="bottom")

p3a

```


Are there differences between buildings, within each sample type and are there tray/cage effects?


```{r}

#############  adults ############### 

adult_uw_dist_matrx = as.matrix(adult_uw_dist)
adult_uw_dist_df <- cbind(rownames(adult_uw_dist_matrx), data.frame(adult_uw_dist_matrx, row.names =NULL))
colnames(adult_uw_dist_df)[1] <- "SampleID"
adult_uw_joined <- inner_join(adult_uw_dist_df, adult_sample_data, by = "SampleID") 

adonis2(formula = adult_uw_dist ~ Building + Building/Tray_Cage, data = adult_uw_joined) 
```

```{r}
pairwise.adonis2(adult_uw_dist ~ Building + Building/Tray_Cage, data = adult_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

adult_uw_disp_building <- betadisper(adult_uw_dist, adult_uw_joined$Building)

adult_uw_perm_disp_building <- permutest(adult_uw_disp_building, permutations = 999)
print(adult_uw_perm_disp_building) # signif

```

```{r}

adult_uw_dists <- list()
adult_uw_sample_data_frames <- list()
adult_uw_dist_matrices <- list()
adult_uw_joined_dfs <- list()
adult_uw_disp_Tray_Cage <- list()
adult_uw_perm_disp_Tray_Cage <- list()

for (i in seq_along(buildings)) {
  current_building <- buildings[i]
  current_label <- labels[i]
  
  adult_uw_subset_data <- subset_samples(ps_adult, Building == current_building)
  
  adult_uw_dist <- phyloseq::distance(adult_uw_subset_data, method="unifrac", weighted=FALSE)
  
  adult_uw_sample_data_df <- as.data.frame(sample_data(adult_uw_subset_data))
  
  adult_uw_dist_matrix <- as.matrix(adult_uw_dist)
  adult_uw_dist_df <- cbind(rownames(adult_uw_dist_matrix), data.frame(adult_uw_dist_matrix, row.names=NULL))
  colnames(adult_uw_dist_df)[1] <- "SampleID"
  colnames(adult_uw_sample_data_df)[1] <- "SampleID"
  
  adult_uw_joined <- inner_join(adult_uw_dist_df, adult_uw_sample_data_df, by="SampleID")
  
  adult_uw_disp_Tray_Cage_result <- betadisper(adult_uw_dist, adult_uw_joined$Tray_Cage)
  adult_uw_perm_disp_Tray_Cage_result <- permutest(adult_uw_disp_Tray_Cage_result, permutations=999)
  
  cat("Results for Building adult UW:", current_building, "\n")
  print(adult_uw_perm_disp_Tray_Cage_result)
  
  adult_uw_dists[[current_label]] <- adult_uw_dist
  adult_uw_sample_data_frames[[current_label]] <- adult_uw_sample_data_df
  adult_uw_dist_matrices[[current_label]] <- adult_uw_dist_matrix
  adult_uw_joined_dfs[[current_label]] <- adult_uw_joined
  adult_uw_disp_Tray_Cage[[current_label]] <- adult_uw_disp_Tray_Cage_result
  adult_uw_perm_disp_Tray_Cage[[current_label]] <- adult_uw_perm_disp_Tray_Cage_result
}

# none signif
```


```{r}

#############  larvae ############### 

larvae_uw_dist_matrx = as.matrix(larvae_uw_dist)
larvae_uw_dist_df <- cbind(rownames(larvae_uw_dist_matrx), data.frame(larvae_uw_dist_matrx, row.names =NULL))
colnames(larvae_uw_dist_df)[1] <- "SampleID"
larvae_uw_joined <- inner_join(larvae_uw_dist_df, larvae_sample_data, by = "SampleID") 

adonis2(formula = larvae_uw_dist ~ Building + Building/Tray_Cage, data = larvae_uw_joined) 

```

```{r}
pairwise.adonis2(larvae_uw_dist ~ Building + Building/Tray_Cage, data = larvae_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

larvae_uw_disp_building <- betadisper(larvae_uw_dist, larvae_uw_joined$Building)

larvae_uw_perm_disp_building <- permutest(larvae_uw_disp_building, permutations = 999)
print(larvae_uw_perm_disp_building)

```


```{r}

larvae_uw_dists <- list()
larvae_uw_sample_data_frames <- list()
larvae_uw_dist_matrices <- list()
larvae_uw_joined_dfs <- list()
larvae_uw_disp_Tray_Cage <- list()
larvae_uw_perm_disp_Tray_Cage <- list()

for (i in seq_along(buildings)) {
  current_building <- buildings[i]
  current_label <- labels[i]
  
  larvae_uw_subset_data <- subset_samples(ps_larvae, Building == current_building)
  
  larvae_uw_dist <- phyloseq::distance(larvae_uw_subset_data, method="unifrac", weighted=FALSE)
  
  larvae_uw_sample_data_df <- as.data.frame(sample_data(larvae_uw_subset_data))
  
  larvae_uw_dist_matrix <- as.matrix(larvae_uw_dist)
  larvae_uw_dist_df <- cbind(rownames(larvae_uw_dist_matrix), data.frame(larvae_uw_dist_matrix, row.names=NULL))
  colnames(larvae_uw_dist_df)[1] <- "SampleID"
  colnames(larvae_uw_sample_data_df)[1] <- "SampleID"
  
  larvae_uw_joined <- inner_join(larvae_uw_dist_df, larvae_uw_sample_data_df, by="SampleID")
  
  larvae_uw_disp_Tray_Cage_result <- betadisper(larvae_uw_dist, larvae_uw_joined$Tray_Cage)
  larvae_uw_perm_disp_Tray_Cage_result <- permutest(larvae_uw_disp_Tray_Cage_result, permutations=999)
  
  cat("Results for Building larvae UW:", current_building, "\n")
  print(larvae_uw_perm_disp_Tray_Cage_result)
  
  larvae_uw_dists[[current_label]] <- larvae_uw_dist
  larvae_uw_sample_data_frames[[current_label]] <- larvae_uw_sample_data_df
  larvae_uw_dist_matrices[[current_label]] <- larvae_uw_dist_matrix
  larvae_uw_joined_dfs[[current_label]] <- larvae_uw_joined
  larvae_uw_disp_Tray_Cage[[current_label]] <- larvae_uw_disp_Tray_Cage_result
  larvae_uw_perm_disp_Tray_Cage[[current_label]] <- larvae_uw_perm_disp_Tray_Cage_result
}

# none signif
```

```{r}

#############  larval water ############### 

larvalwater_uw_dist_matrx = as.matrix(larvalwater_uw_dist)
larvalwater_uw_dist_df <- cbind(rownames(larvalwater_uw_dist_matrx), data.frame(larvalwater_uw_dist_matrx, row.names =NULL))
colnames(larvalwater_uw_dist_df)[1] <- "SampleID"
larvalwater_uw_joined <- inner_join(larvalwater_uw_dist_df, larvalwater_sample_data, by = "SampleID") 

adonis2(formula = larvalwater_uw_dist ~ Building + Building/Tray_Cage, data = larvalwater_uw_joined) 

```

```{r}
pairwise.adonis2(larvalwater_uw_dist ~ Building + Building/Tray_Cage, data = larvalwater_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

larvalwater_uw_disp_building <- betadisper(larvalwater_uw_dist, larvalwater_uw_joined$Building)

larvalwater_uw_perm_disp_building <- permutest(larvalwater_uw_disp_building, permutations = 999)
print(larvalwater_uw_perm_disp_building)

```


```{r}

LW_uw_dists <- list()
LW_uw_sample_data_frames <- list()
LW_uw_dist_matrices <- list()
LW_uw_joined_dfs <- list()
LW_uw_disp_Tray_Cage <- list()
LW_uw_perm_disp_Tray_Cage <- list()

for (i in seq_along(buildings)) {
  current_building <- buildings[i]
  current_label <- labels[i]
  
  LW_uw_subset_data <- subset_samples(ps_larvalwater, Building == current_building)
  
  LW_uw_dist <- phyloseq::distance(LW_uw_subset_data, method="unifrac", weighted=FALSE)
  
  LW_uw_sample_data_df <- as.data.frame(sample_data(LW_uw_subset_data))
  
  LW_uw_dist_matrix <- as.matrix(LW_uw_dist)
  LW_uw_dist_df <- cbind(rownames(LW_uw_dist_matrix), data.frame(LW_uw_dist_matrix, row.names=NULL))
  colnames(LW_uw_dist_df)[1] <- "SampleID"
  colnames(LW_uw_sample_data_df)[1] <- "SampleID"
  
  LW_uw_joined <- inner_join(LW_uw_dist_df, LW_uw_sample_data_df, by="SampleID")
  
  LW_uw_disp_Tray_Cage_result <- betadisper(LW_uw_dist, LW_uw_joined$Tray_Cage)
  LW_uw_perm_disp_Tray_Cage_result <- permutest(LW_uw_disp_Tray_Cage_result, permutations=999)
  
  cat("Results for Building larval water UW:", current_building, "\n")
  print(LW_uw_perm_disp_Tray_Cage_result)
  
  LW_uw_dists[[current_label]] <- LW_uw_dist
  LW_uw_sample_data_frames[[current_label]] <- LW_uw_sample_data_df
  LW_uw_dist_matrices[[current_label]] <- LW_uw_dist_matrix
  LW_uw_joined_dfs[[current_label]] <- LW_uw_joined
  LW_uw_disp_Tray_Cage[[current_label]] <- LW_uw_disp_Tray_Cage_result
  LW_uw_perm_disp_Tray_Cage[[current_label]] <- LW_uw_perm_disp_Tray_Cage_result
}

# none signif
```

```{r}

#############  tap water ############### 

tapwater_uw_dist_matrx = as.matrix(water_uw_dist)
tapwater_uw_dist_df <- cbind(rownames(tapwater_uw_dist_matrx), data.frame(tapwater_uw_dist_matrx, row.names =NULL))
colnames(tapwater_uw_dist_df)[1] <- "SampleID"
tapwater_uw_joined <- inner_join(tapwater_uw_dist_df, tapwater_sample_data, by = "SampleID") 

adonis2(formula = water_uw_dist ~ Building, data = tapwater_uw_joined) # no tray effects tested for with tap water

```

```{r}
pairwise.adonis2(water_uw_dist ~ Building, data = tapwater_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```


```{r}

tapwater_uw_disp_building <- betadisper(water_uw_dist, tapwater_uw_joined$Building)

tapwater_uw_perm_disp_building <- permutest(tapwater_uw_disp_building, permutations = 999)
print(tapwater_uw_perm_disp_building)

```


```{r}

#############  fish food  ############### 

food_uw_dist_matrx = as.matrix(food_uw_dist)
food_uw_dist_df <- cbind(rownames(food_uw_dist_matrx), data.frame(food_uw_dist_matrx, row.names =NULL))
colnames(food_uw_dist_df)[1] <- "SampleID"
food_uw_joined <- inner_join(food_uw_dist_df, food_sample_data, by = "SampleID") 

adonis2(formula = food_uw_dist ~ Building, data = food_uw_joined) # no tray effects tested for with food

```
not statistically significant

```{r}

food_uw_disp_building <- betadisper(food_uw_dist, food_uw_joined$Building)

food_uw_perm_disp_building <- permutest(food_uw_disp_building, permutations = 999)
print(food_uw_perm_disp_building)

```

```{r}

#############  sugar solution ############### 

sugar_uw_dist_matrx = as.matrix(sugar_uw_dist)
sugar_uw_dist_df <- cbind(rownames(sugar_uw_dist_matrx), data.frame(sugar_uw_dist_matrx, row.names =NULL))
colnames(sugar_uw_dist_df)[1] <- "SampleID"
sugar_uw_joined <- inner_join(sugar_uw_dist_df, sugar_sample_data, by = "SampleID") 

adonis2(formula = sugar_uw_dist ~ Building, data = sugar_uw_joined) 

```

```{r}
pairwise.adonis2(sugar_uw_dist ~ Building, data = sugar_uw_joined, p.adjust.method = "bonferroni", nperm = 999)

```

```{r}

sugar_uw_disp_building <- betadisper(sugar_uw_dist, sugar_uw_joined$Building)

sugar_uw_perm_disp_building <- permutest(sugar_uw_disp_building, permutations = 999)
print(sugar_uw_perm_disp_building)

```


