---
title: "Ins_comp_8-TTsum"
output: html_document
date: "2024-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TinyTag summaries

TinyTag data. Copied raw values into text document via vi as xls sheet has formatting.

```{r libraries_cols}
library(ggplot2)
library(tidyverse)
library(patchwork)
library(onewaytests)

colours2 <- c(
  "C" = "#00A087FF",
  "A" = "#3C5488FF",
  "B" = "#9C58A1"
)

tinytag <- read.table("TT_data.tsv", header=T)

```

## Plots

Trying to show summaries to better illustratey how different especially insectaries A and B are from C in terms of stability of environmental factors.

```{r plots_boxplot}

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}

p <- ggplot(tinytag, aes(x=insectary, y=Temperature, fill=insectary)) +
  geom_violin() +
  scale_fill_manual(values=colours2) +
  stat_summary(fun.data=data_summary) +
  theme_minimal() +
  labs(title="Temperature measurements",x="Insectary", y = "Temperature (C)")

p2 <- ggplot(tinytag, aes(x=insectary, y=Humidity, fill=insectary)) +
  geom_violin() +
  scale_fill_manual(values=colours2) +
  stat_summary(fun.data=data_summary) +
  theme_minimal() +
  labs(title="Humidity measurements",x="Insectary", y = "Humidity (% RH)")


p3 <- ggplot(tinytag, aes(x=insectary, y=Temperature, fill=insectary)) +
  geom_violin() +
  scale_fill_manual(values=colours2) +
  geom_boxplot(width=0.1, fill="white", alpha=0.3) +
  theme_minimal() +
  labs(title="Temperature measurements",x="Insectary", y = "Temperature (C)")

p4 <- ggplot(tinytag, aes(x=insectary, y=Humidity, fill=insectary)) +
  geom_violin() +
  scale_fill_manual(values=colours2) +
  geom_boxplot(width=0.1, fill="white", alpha=0.3) +
  theme_minimal() +
  labs(title="Humidity measurements",x="Insectary", y = "Humidity (% RH)")

p3 + p4 + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') +  plot_annotation(title = 'TinyTag data')

```



Taking a look at the data spread (not difference between values altogether/means as these were set differently, so obvioulsy different)

```{r}

tinytag %>%
     group_by(insectary) %>%
     summarize(var=var(Temperature))

bf.test(Temperature ~ insectary, data = tinytag)


tinytag %>%
     group_by(insectary) %>%
     summarize(var=var(Humidity))

bf.test(Humidity ~ insectary, data = tinytag)

```

