---
title: "Ins_comp_7-survival"
author: "laura"
date: "2024-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.


# Install packages

```{r}
library(ggplot2)
library(patchwork)
library(tidyverse)
```

# import data

```{r}
dat = read.csv("ins_comp_survival_data.csv")


sapply(dat, class)

as.numeric(dat$number)


```

```{r}
colours2 <- c(
  "ACC" = "#00A087FF",
  "CTID" = "#3C5488FF",
  "MK" = "#9C58A1"
)
```

# plot 

```{r}

p <- ggplot(data= dat, aes(x= tray, y= number, colour = insectary, shape = stage)) +
  geom_point(size = 3) + 
   scale_color_manual(values = colours2) +
  scale_y_continuous(limits = c(90, 150)) +
  facet_grid(~factor(insectary, levels=c("CTID", "MK", "ACC")), scales = "free", space = "free")
p


```

# stats

using the same data presented in a slightly different way


```{r}

# import data
dat2 = read.csv("ins_comp_survival2.csv")

# sapply(dat2, class)

```

```{r}

# ANOVA
anova_pupae <- aov(stage2 ~ insectary, data = dat2)
anova_adult <- aov(stage3 ~ insectary, data = dat2)

summary(anova_pupae)
summary(anova_adult)


# Kruskal-WAllis
Kruskal_pupae <- kruskal.test(stage2 ~ insectary, data = dat2)
Kruskal_adult <- kruskal.test(stage3 ~ insectary, data = dat2)


if (Kruskal_pupae$p.value < 0.05) {
  print("There are significant differences between at least two groups.")
} else {
  print("There is not enough evidence to conclude significant differences between groups.")
}

if (Kruskal_adult$p.value < 0.05) {
  print("There are significant differences between at least two groups.")
} else {
  print("There is not enough evidence to conclude significant differences between groups.")
}


```

No statistically significant difference in numbers making it to pupal and adult stages in the different rooms, but only 3 reps per room so it would need to be a big difference to see that.


# pupal development time


```{r}
pup = read.csv("pupae_development.csv")

```

# plot 

```{r}

p2 <- ggplot(data= pup, aes(x= time.to.pupation, y= number, colour = insectary, shape = tray)) +
  geom_point(size = 3) + 
   scale_color_manual(values = colours2) +
  scale_y_continuous(limits = c(0, 150)) +
  facet_grid(~factor(insectary, levels=c("CTID", "MK", "ACC")), scales = "free", space = "free")+
  geom_line()
p2


#Renaming insectaries to make my life easier

pup <- pup %>% 
  mutate(insectary = gsub("CTID", "A", insectary), 
         insectary = gsub("MK", "B", insectary),
         insectary = gsub("ACC", "C", insectary)
  )

dat <- dat %>% 
  mutate(insectary = gsub("CTID", "A", insectary), 
         insectary = gsub("MK", "B", insectary),
         insectary = gsub("ACC", "C", insectary)
  )


dat <- dat %>% 
  mutate(tray = gsub("A", "t1", tray), 
         tray = gsub("B", "t2", tray),
         tray = gsub("C", "t3", tray)
  )



colours3 <- c(
  "C" = "#00A087FF",
  "A" = "#3C5488FF",
  "B" = "#9C58A1"
)


p3 <- ggplot(data = pup, mapping = aes(time.to.pupation, number, fill=insectary)) + 
  geom_boxplot(alpha=0.8) +
  scale_fill_manual(values = colours3) +
  theme_minimal() +
  geom_point(position=position_dodge(width=0.75),aes(group=insectary)) + 
  labs(title="Pupation time",x="Day of experiment", y = "Count (mosquitoes pupated)")


p <- ggplot(data= dat, aes(x= tray, y= number, colour = insectary, shape = stage)) +
  geom_point(size = 2) + 
  scale_color_manual(values = colours3) +
  scale_y_continuous(limits = c(90, 150)) +
  facet_grid(~factor(insectary, levels=c("A", "B", "C")), scales = "free", space = "free") +
  theme_bw() + 
  labs(title="Developmental stages",x="Tray", y = "Count (mosquitoes)")

p3 / p + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') +  plot_annotation(title = 'Mosquito development times')

pdf("Mosquito_development.pdf", width=4, height=7)
p3 / p + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') +  plot_annotation(title = 'Mosquito development times')
dev.off()
```

