---
title: "Ins_comp_6a-Dif_adbund"
author: "laura"
date: "2024-01-17"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA, 
                      fig.width = 6.25, fig.height = 5)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.


Now  to investigate whether there are particular taxa are associated with different insectaries.


This paper 'https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-017-0237-y' suggests ANCOM is better than DESeq2 for micro data 'For drawing inferences regarding taxon abundance in the ecosystem, analysis of composition of microbiomes (ANCOM) is not only very sensitive (for >20 samples per group) but also critically the only method tested that has a good control of false discovery rate.'

REcently, ANCOM-BC2 has been developed to improve upon ANCOM-BC. Here I am using info from the following vignette:

https://www.bioconductor.org/packages/release/bioc/vignettes/ANCOMBC/inst/doc/ANCOMBC2.html 

with info from the ANCOM-BC-2 paper

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10776411/


# Install packages

```{r}


library(DT)
options(DT.options = list(
  initComplete = JS("function(settings, json) {",
  "$(this.api().table().header()).css({'background-color': 
  '#000', 'color': '#fff'});","}")))

# It appears to be a package compatibility issue between the release version of 
# phyloseq and lme4, a fresh installation of phyloseq might be needed
# See this post: https://github.com/lme4/lme4/issues/743
# remotes::install_github("joey711/phyloseq", force = TRUE)

library(phyloseq)
library(ggplot2)
library(tidyverse)
library(ANCOMBC)

```

# import data

```{r}

rm(list = ls()) 
ps = readRDS("physeq_obj_3974") 
head(sample_data(ps))

```

4.2 Run ancombc2 function
To control the FDR arising from multiple testing, we opt for the Holm-Bonferroni method over the Benjamini-Hochberg (BH) procedure, especially when dealing with large sample sizes where statistical power isn’t the primary concern. The Holm-Bonferroni method, accommodating any dependence structure among p-values, is known to be robust against inaccuracies in p-values, an issue often seen in DA analysis. Figures below display only results significant after the Holm-Bonferroni adjustment.

```{r, warning=FALSE}
set.seed(123)

output = ancombc2(data = ps, assay_name = "counts", tax_level = "Genus",
                  fix_formula = "Sample_Type2 + Building", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, s0_perc = 0.05,
                  group = "Building", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, dunnet = FALSE, trend = FALSE,
                  iter_control = list(tol = 1e-2, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(-1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2, 2, 1),
                                       solver = "ECOS",
                                       B = 10))


```


4.3 Structural zeros (taxon presence/absence)

```{r}
tab_zero = output$zero_ind
tab_zero %>%
    datatable(caption = "The detection of structural zeros")

```

as expected there are many taxa with structural zeros


4.4 ANCOM-BC2 primary analysis
The primary output of the ANCOM-BC2 methodology identifies taxa with differential abundance based on the chosen covariate. The results include: 1) log fold changes, 2) standard errors, 3) test statistics, 4) p-values, 5) adjusted p-values, 6) indicators denoting whether the taxon is differentially abundant (TRUE) or not (FALSE), and 7) indicators denoting whether the taxon passed the sensitivity analysis (TRUE) or not (FALSE).


```{r}

res_prim = output$res

```



4.6 ANCOM-BC2 multiple pairwise comparisons
The ANCOM-BC2 methodology for multiple pairwise comparisons is designed to identify taxa that exhibit differential abundance between any two groups within a set of three or more experimental groups, all while maintaining control over the mdFDR.

For instance, in our analysis focusing on the categories “lean”, “overweight”, and “obese”, the output provides: 1) log fold changes, 2) standard errors, 3) test statistics, 4) p-values, 5) adjusted p-values, 6) indicators denoting whether the taxon is differentially abundant (TRUE) or not (FALSE), and 7) indicators denoting whether the taxon passed the sensitivity analysis (TRUE) or not (FALSE).

In the subsequent heatmap, each cell represents a log fold-change (in natural log) value. Entries highlighted in green have successfully passed the sensitivity analysis for pseudo-count addition.


```{r}
res_pair = output$res_pair

colnames(res_pair) <- gsub("-", "", colnames(res_pair)) 

df_fig_pair1 = res_pair %>%
    dplyr::filter(diff_BuildingBMK == 1 |
                      diff_BuildingCACC == 1 | 
                      diff_BuildingCACC_BuildingBMK == 1) %>%
    dplyr::mutate(lfc1 = ifelse(diff_BuildingBMK == 1, 
                                round(lfc_BuildingBMK, 2), 0),
                  lfc2 = ifelse(diff_BuildingCACC == 1, 
                                round(lfc_BuildingCACC, 2), 0),
                  lfc3 = ifelse(diff_BuildingCACC_BuildingBMK == 1, 
                                round(lfc_BuildingCACC_BuildingBMK, 2), 0)) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "value") %>%
    dplyr::arrange(taxon)

df_fig_pair2 = res_pair %>%
    dplyr::filter(diff_BuildingBMK == 1 |
                      diff_BuildingCACC == 1 | 
                      diff_BuildingCACC_BuildingBMK == 1) %>%
    dplyr::mutate(lfc1 = ifelse(passed_ss_BuildingBMK == 1 & diff_BuildingBMK == 1, 
                                "#00A087FF", "black"),
                  lfc2 = ifelse(passed_ss_BuildingCACC == 1 & diff_BuildingCACC == 1, 
                                "#00A087FF", "black"),
                  lfc3 = ifelse(passed_ss_BuildingCACC_BuildingBMK == 1 & diff_BuildingCACC_BuildingBMK == 1, 
                                "#00A087FF", "black")) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "color") %>%
    dplyr::arrange(taxon)

df_fig_pair = df_fig_pair1 %>%
    dplyr::left_join(df_fig_pair2, by = c("taxon", "group"))

df_fig_pair$group = recode(df_fig_pair$group, 
                          `lfc1` = "Ins_B - Ins_A",
                          `lfc2` = "Ins_C - Ins_A",
                          `lfc3` = "Ins_C - Ins_B")

df_fig_pair$group = factor(df_fig_pair$group, 
                          levels = c("Ins_B - Ins_A",
                                     "Ins_C - Ins_A", 
                                     "Ins_C - Ins_B"))

lo = floor(min(df_fig_pair$value))
up = ceiling(max(df_fig_pair$value))
mid = (lo + up)/2
fig_pair = df_fig_pair %>%
    ggplot(aes(x = group, y = taxon, fill = value)) + 
    geom_tile(color = "black") +
    scale_fill_gradient2(low = "#002F6C", high = "#AC0123", mid = "white", 
                         na.value = "white", midpoint = 0, limit = c(lo, up),
                         name = NULL) +
    geom_text(aes(group, taxon, label = value, color = color), size = 4) +
    scale_color_identity(guide = FALSE) +
    labs(x = NULL, y = NULL, title = NULL) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
fig_pair

```

# now for tray/cage differences

Here I will separate by building and sample type. I will keep larvae and larval water together as they are literally the same tray and separate cage as the associated ASVs may be different. It won't let me set the formula to nested effect of building/tray_cage so will sub sample by building too.



## larvae and larval water (tray), insectary A

```{r, warning=FALSE}
ps_A_tray <- subset_samples(ps, Sample_Type2 %in% c("Larva", "LarvalWater"))
ps_A_tray <- subset_samples(ps_A_tray, Building == "A-CTID")


set.seed(123)

output_A_tray = ancombc2(data = ps_A_tray, assay_name = "counts", tax_level = "Genus",
                  fix_formula = "Tray_Cage", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, s0_perc = 0.05,
                  group = "Tray_Cage", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = FALSE,
                  global = TRUE, pairwise = TRUE, dunnet = FALSE, trend = FALSE,
                  iter_control = list(tol = 1e-2, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(-1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2, 2, 1),
                                       solver = "ECOS",
                                       B = 10))

```

```{r}
res_pair_AT = output_A_tray$res_pair

colnames(res_pair_AT) <- gsub("-", "", colnames(res_pair_AT)) 

df_fig_pair1_AT = res_pair_AT %>%
    dplyr::filter(diff_Tray_CageTray2 == 1 |
                      diff_Tray_CageTray3 == 1 | 
                      diff_Tray_CageTray3_Tray_CageTray2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(diff_Tray_CageTray2 == 1, 
                                round(lfc_Tray_CageTray2, 2), 0),
                  lfc2 = ifelse(diff_Tray_CageTray3 == 1, 
                                round(lfc_Tray_CageTray3, 2), 0),
                  lfc3 = ifelse(diff_Tray_CageTray3_Tray_CageTray2 == 1, 
                                round(lfc_Tray_CageTray3_Tray_CageTray2, 2), 0)) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "value") %>%
    dplyr::arrange(taxon)

df_fig_pair2_AT = res_pair_AT %>%
    dplyr::filter(diff_Tray_CageTray2 == 1 |
                      diff_Tray_CageTray3 == 1 | 
                      diff_Tray_CageTray3_Tray_CageTray2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(passed_ss_Tray_CageTray2 == 1 & diff_Tray_CageTray2 == 1, 
                                "#00A087FF", "black"),
                  lfc2 = ifelse(passed_ss_Tray_CageTray3 == 1 & diff_Tray_CageTray3 == 1, 
                                "#00A087FF", "black"),
                  lfc3 = ifelse(passed_ss_Tray_CageTray3_Tray_CageTray2 == 1 & diff_Tray_CageTray3_Tray_CageTray2 == 1, 
                                "#00A087FF", "black")) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "color") %>%
    dplyr::arrange(taxon)

df_fig_pair_AT = df_fig_pair1_AT %>%
    dplyr::left_join(df_fig_pair2_AT, by = c("taxon", "group"))

df_fig_pair_AT$group = recode(df_fig_pair_AT$group, 
                          `lfc1` = "tray_2 - tray_1",
                          `lfc2` = "tray_3 - tray_1",
                          `lfc3` = "tray_3 - tray_2")

df_fig_pair_AT$group = factor(df_fig_pair_AT$group, 
                          levels = c("tray_2 - tray_1",
                                     "tray_3 - tray_1", 
                                     "tray_3 - tray_2"))


fig_pair_AT = df_fig_pair_AT %>%
    ggplot(aes(x = group, y = taxon, fill = value)) + 
    geom_tile(color = "black") +
    scale_fill_gradient2(low = "#002F6C", high = "#AC0123", mid = "white", 
                         na.value = "white", midpoint = 0, limit = c(-4, 4),
                         name = NULL) +
    geom_text(aes(group, taxon, label = value, color = color), size = 4) +
    scale_color_identity(guide = FALSE) +
    labs(x = NULL, y = NULL, title = "tray_insA") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
fig_pair_AT

```


## larvae and larval water (tray), insectary B

```{r, warning=FALSE}
ps_B_tray <- subset_samples(ps, Sample_Type2 %in% c("Larva", "LarvalWater"))
ps_B_tray <- subset_samples(ps_B_tray, Building == "B-MK")


set.seed(123)

output_B_tray = ancombc2(data = ps_B_tray, assay_name = "counts", tax_level = "Genus",
                  fix_formula = "Tray_Cage", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, s0_perc = 0.05,
                  group = "Tray_Cage", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = FALSE,
                  global = TRUE, pairwise = TRUE, dunnet = FALSE, trend = FALSE,
                  iter_control = list(tol = 1e-2, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(-1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2, 2, 1),
                                       solver = "ECOS",
                                       B = 10))

```


```{r}
res_pair_BT = output_B_tray$res_pair

colnames(res_pair_BT) <- gsub("-", "", colnames(res_pair_BT)) 

df_fig_pair1_BT = res_pair_BT %>%
    dplyr::filter(diff_Tray_CageTray2 == 1 |
                      diff_Tray_CageTray3 == 1 | 
                      diff_Tray_CageTray3_Tray_CageTray2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(diff_Tray_CageTray2 == 1, 
                                round(lfc_Tray_CageTray2, 2), 0),
                  lfc2 = ifelse(diff_Tray_CageTray3 == 1, 
                                round(lfc_Tray_CageTray3, 2), 0),
                  lfc3 = ifelse(diff_Tray_CageTray3_Tray_CageTray2 == 1, 
                                round(lfc_Tray_CageTray3_Tray_CageTray2, 2), 0)) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "value") %>%
    dplyr::arrange(taxon)

df_fig_pair2_BT = res_pair_BT %>%
    dplyr::filter(diff_Tray_CageTray2 == 1 |
                      diff_Tray_CageTray3 == 1 | 
                      diff_Tray_CageTray3_Tray_CageTray2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(passed_ss_Tray_CageTray2 == 1 & diff_Tray_CageTray2 == 1, 
                                "#00A087FF", "black"),
                  lfc2 = ifelse(passed_ss_Tray_CageTray3 == 1 & diff_Tray_CageTray3 == 1, 
                                "#00A087FF", "black"),
                  lfc3 = ifelse(passed_ss_Tray_CageTray3_Tray_CageTray2 == 1 & diff_Tray_CageTray3_Tray_CageTray2 == 1, 
                                "#00A087FF", "black")) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "color") %>%
    dplyr::arrange(taxon)

df_fig_pair_BT = df_fig_pair1_BT %>%
    dplyr::left_join(df_fig_pair2_BT, by = c("taxon", "group"))

df_fig_pair_BT$group = recode(df_fig_pair_BT$group, 
                          `lfc1` = "tray_2 - tray_1",
                          `lfc2` = "tray_3 - tray_1",
                          `lfc3` = "tray_3 - tray_2")

df_fig_pair_BT$group = factor(df_fig_pair_BT$group, 
                          levels = c("tray_2 - tray_1",
                                     "tray_3 - tray_1", 
                                     "tray_3 - tray_2"))


fig_pair_BT = df_fig_pair_BT %>%
    ggplot(aes(x = group, y = taxon, fill = value)) + 
    geom_tile(color = "black") +
    scale_fill_gradient2(low = "#002F6C", high = "#AC0123", mid = "white", 
                         na.value = "white", midpoint = 0, limit = c(-4, 4),
                         name = NULL) +
    geom_text(aes(group, taxon, label = value, color = color), size = 4) +
    scale_color_identity(guide = FALSE) +
    labs(x = NULL, y = NULL, title = "tray_insB") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
fig_pair_BT

```

## larvae and larval water (tray), insectary C

```{r, warning=FALSE}
ps_C_tray <- subset_samples(ps, Sample_Type2 %in% c("Larva", "LarvalWater"))
ps_C_tray <- subset_samples(ps_C_tray, Building == "C-ACC")


set.seed(123)

output_C_tray = ancombc2(data = ps_C_tray, assay_name = "counts", tax_level = "Genus",
                  fix_formula = "Tray_Cage", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, s0_perc = 0.05,
                  group = "Tray_Cage", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, dunnet = FALSE, trend = FALSE,
                  iter_control = list(tol = 1e-2, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(-1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2, 2, 1),
                                       solver = "ECOS",
                                       B = 10))

```

```{r}
res_pair_CT = output_C_tray$res_pair

colnames(res_pair_CT) <- gsub("-", "", colnames(res_pair_CT)) 

df_fig_pair1_CT = res_pair_CT %>%
    dplyr::filter(diff_Tray_CageTray2 == 1 |
                      diff_Tray_CageTray3 == 1 | 
                      diff_Tray_CageTray3_Tray_CageTray2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(diff_Tray_CageTray2 == 1, 
                                round(lfc_Tray_CageTray2, 2), 0),
                  lfc2 = ifelse(diff_Tray_CageTray3 == 1, 
                                round(lfc_Tray_CageTray3, 2), 0),
                  lfc3 = ifelse(diff_Tray_CageTray3_Tray_CageTray2 == 1, 
                                round(lfc_Tray_CageTray3_Tray_CageTray2, 2), 0)) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "value") %>%
    dplyr::arrange(taxon)

df_fig_pair2_CT = res_pair_CT %>%
    dplyr::filter(diff_Tray_CageTray2 == 1 |
                      diff_Tray_CageTray3 == 1 | 
                      diff_Tray_CageTray3_Tray_CageTray2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(passed_ss_Tray_CageTray2 == 1 & diff_Tray_CageTray2 == 1, 
                                "#00A087FF", "black"),
                  lfc2 = ifelse(passed_ss_Tray_CageTray3 == 1 & diff_Tray_CageTray3 == 1, 
                                "#00A087FF", "black"),
                  lfc3 = ifelse(passed_ss_Tray_CageTray3_Tray_CageTray2 == 1 & diff_Tray_CageTray3_Tray_CageTray2 == 1, 
                                "#00A087FF", "black")) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "color") %>%
    dplyr::arrange(taxon)

df_fig_pair_CT = df_fig_pair1_CT %>%
    dplyr::left_join(df_fig_pair2_CT, by = c("taxon", "group"))

df_fig_pair_CT$group = recode(df_fig_pair_CT$group, 
                          `lfc1` = "tray_2 - tray_1",
                          `lfc2` = "tray_3 - tray_1",
                          `lfc3` = "tray_3 - tray_2")

df_fig_pair_CT$group = factor(df_fig_pair_CT$group, 
                          levels = c("tray_2 - tray_1",
                                     "tray_3 - tray_1", 
                                     "tray_3 - tray_2"))


fig_pair_CT = df_fig_pair_CT %>%
    ggplot(aes(x = group, y = taxon, fill = value)) + 
    geom_tile(color = "black") +
    scale_fill_gradient2(low = "#002F6C", high = "#AC0123", mid = "white", 
                         na.value = "white", midpoint = 0, limit = c(-4,4),
                         name = NULL) +
    geom_text(aes(group, taxon, label = value, color = color), size = 4) +
    scale_color_identity(guide = FALSE) +
    labs(x = NULL, y = NULL, title = "tray_insC") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
fig_pair_CT

```




## adult females (cage), insectary A

```{r, warning=FALSE}
ps_A_cage <- subset_samples(ps, Sample_Type2 == "AdultFemale")
ps_A_cage <- subset_samples(ps_A_cage, Building == "A-CTID")


set.seed(123)

output_A_cage = ancombc2(data = ps_A_cage, assay_name = "counts", tax_level = "Genus",
                  fix_formula = "Tray_Cage", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, s0_perc = 0.05,
                  group = "Tray_Cage", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, dunnet = FALSE, trend = FALSE,
                  iter_control = list(tol = 1e-2, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(-1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2, 2, 1),
                                       solver = "ECOS",
                                       B = 10))

```

```{r}
res_pair_AC = output_A_cage$res_pair

colnames(res_pair_AC) <- gsub("-", "", colnames(res_pair_AC)) 

df_fig_pair1_AC = res_pair_AC %>%
    dplyr::filter(diff_Tray_CageCage2 == 1 |
                      diff_Tray_CageCage3 == 1 | 
                      diff_Tray_CageCage3_Tray_CageCage2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(diff_Tray_CageCage2 == 1, 
                                round(lfc_Tray_CageCage2, 2), 0),
                  lfc2 = ifelse(diff_Tray_CageCage3 == 1, 
                                round(lfc_Tray_CageCage3, 2), 0),
                  lfc3 = ifelse(diff_Tray_CageCage3_Tray_CageCage2 == 1, 
                                round(lfc_Tray_CageCage3_Tray_CageCage2, 2), 0)) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "value") %>%
    dplyr::arrange(taxon)

df_fig_pair2_AC = res_pair_AC %>%
    dplyr::filter(diff_Tray_CageCage2 == 1 |
                      diff_Tray_CageCage3 == 1 | 
                      diff_Tray_CageCage3_Tray_CageCage2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(passed_ss_Tray_CageCage2 == 1 & diff_Tray_CageCage2 == 1, 
                                "#00A087FF", "black"),
                  lfc2 = ifelse(passed_ss_Tray_CageCage3 == 1 & diff_Tray_CageCage3 == 1, 
                                "#00A087FF", "black"),
                  lfc3 = ifelse(passed_ss_Tray_CageCage3_Tray_CageCage2 == 1 & diff_Tray_CageCage3_Tray_CageCage2 == 1, 
                                "#00A087FF", "black")) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "color") %>%
    dplyr::arrange(taxon)

df_fig_pair_AC = df_fig_pair1_AC %>%
    dplyr::left_join(df_fig_pair2_AC, by = c("taxon", "group"))

df_fig_pair_AC$group = recode(df_fig_pair_AC$group, 
                          `lfc1` = "cage_2 - cage_1",
                          `lfc2` = "cage_3 - cage_1",
                          `lfc3` = "cage_3 - cage_2")

df_fig_pair_AC$group = factor(df_fig_pair_AC$group, 
                          levels = c("cage_2 - cage_1",
                                     "cage_3 - cage_1", 
                                     "cage_3 - cage_2"))


fig_pair_AC = df_fig_pair_AC %>%
    ggplot(aes(x = group, y = taxon, fill = value)) + 
    geom_tile(color = "black") +
    scale_fill_gradient2(low = "#002F6C", high = "#AC0123", mid = "white", 
                         na.value = "white", midpoint = 0, limit = c(-4, 4),
                         name = NULL) +
    geom_text(aes(group, taxon, label = value, color = color), size = 4) +
    scale_color_identity(guide = FALSE) +
    labs(x = NULL, y = NULL, title = "cage_insA") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
fig_pair_AC

```


## adult females (cage), insectary B

```{r, warning=FALSE}
ps_B_cage <- subset_samples(ps, Sample_Type2 == "AdultFemale")
ps_B_cage <- subset_samples(ps_B_cage, Building == "B-MK")


set.seed(123)

output_B_cage = ancombc2(data = ps_B_cage, assay_name = "counts", tax_level = "Genus",
                  fix_formula = "Tray_Cage", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, s0_perc = 0.05,
                  group = "Tray_Cage", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, dunnet = FALSE, trend = FALSE,
                  iter_control = list(tol = 1e-2, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(-1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2, 2, 1),
                                       solver = "ECOS",
                                       B = 10))

```


```{r}
res_pair_BC = output_B_cage$res_pair

colnames(res_pair_BC) <- gsub("-", "", colnames(res_pair_BC)) 

df_fig_pair1_BC = res_pair_BC %>%
    dplyr::filter(diff_Tray_CageCage2 == 1 |
                      diff_Tray_CageCage3 == 1 | 
                      diff_Tray_CageCage3_Tray_CageCage2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(diff_Tray_CageCage2 == 1, 
                                round(lfc_Tray_CageCage2, 2), 0),
                  lfc2 = ifelse(diff_Tray_CageCage3 == 1, 
                                round(lfc_Tray_CageCage3, 2), 0),
                  lfc3 = ifelse(diff_Tray_CageCage3_Tray_CageCage2 == 1, 
                                round(lfc_Tray_CageCage3_Tray_CageCage2, 2), 0)) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "value") %>%
    dplyr::arrange(taxon)

df_fig_pair2_BC = res_pair_BC %>%
    dplyr::filter(diff_Tray_CageCage2 == 1 |
                      diff_Tray_CageCage3 == 1 | 
                      diff_Tray_CageCage3_Tray_CageCage2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(passed_ss_Tray_CageCage2 == 1 & diff_Tray_CageCage2 == 1, 
                                "#00A087FF", "black"),
                  lfc2 = ifelse(passed_ss_Tray_CageCage3 == 1 & diff_Tray_CageCage3 == 1, 
                                "#00A087FF", "black"),
                  lfc3 = ifelse(passed_ss_Tray_CageCage3_Tray_CageCage2 == 1 & diff_Tray_CageCage3_Tray_CageCage2 == 1, 
                                "#00A087FF", "black")) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "color") %>%
    dplyr::arrange(taxon)

df_fig_pair_BC = df_fig_pair1_BC %>%
    dplyr::left_join(df_fig_pair2_BC, by = c("taxon", "group"))

df_fig_pair_BC$group = recode(df_fig_pair_BC$group, 
                          `lfc1` = "cage_2 - cage_1",
                          `lfc2` = "cage_3 - cage_1",
                          `lfc3` = "cage_3 - cage_2")

df_fig_pair_BC$group = factor(df_fig_pair_BC$group, 
                          levels = c("cage_2 - cage_1",
                                     "cage_3 - cage_1", 
                                     "cage_3 - cage_2"))

fig_pair_BC = df_fig_pair_BC %>%
    ggplot(aes(x = group, y = taxon, fill = value)) + 
    geom_tile(color = "black") +
    scale_fill_gradient2(low = "#002F6C", high = "#AC0123", mid = "white", 
                         na.value = "white", midpoint = 0, limit = c(-4, 4),
                         name = NULL) +
    geom_text(aes(group, taxon, label = value, color = color), size = 4) +
    scale_color_identity(guide = FALSE) +
    labs(x = NULL, y = NULL, title = "cage_insB") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
fig_pair_BC

```


## adult females (cage), insectary C

```{r, warning=FALSE}
ps_C_cage <- subset_samples(ps, Sample_Type2 == "AdultFemale")
ps_C_cage <- subset_samples(ps_C_cage, Building == "C-ACC")


set.seed(123)

output_C_cage = ancombc2(data = ps_C_cage, assay_name = "counts", tax_level = "Genus",
                  fix_formula = "Tray_Cage", rand_formula = NULL,
                  p_adj_method = "holm", pseudo_sens = TRUE,
                  prv_cut = 0.10, s0_perc = 0.05,
                  group = "Tray_Cage", struc_zero = TRUE, neg_lb = TRUE,
                  alpha = 0.05, n_cl = 2, verbose = TRUE,
                  global = TRUE, pairwise = TRUE, dunnet = FALSE, trend = FALSE,
                  iter_control = list(tol = 1e-2, max_iter = 20, 
                                      verbose = TRUE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = lme4::lmerControl(),
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100),
                  trend_control = list(contrast = list(matrix(c(1, 0, -1, 1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(-1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE),
                                                       matrix(c(1, 0, 1, -1),
                                                              nrow = 2, 
                                                              byrow = TRUE)),
                                       node = list(2, 2, 1),
                                       solver = "ECOS",
                                       B = 10))

```

```{r}
res_pair_CC = output_C_cage$res_pair

colnames(res_pair_CC) <- gsub("-", "", colnames(res_pair_CC)) 

df_fig_pair1_CC = res_pair_CC %>%
    dplyr::filter(diff_Tray_CageCage2 == 1 |
                      diff_Tray_CageCage3 == 1 | 
                      diff_Tray_CageCage3_Tray_CageCage2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(diff_Tray_CageCage2 == 1, 
                                round(lfc_Tray_CageCage2, 2), 0),
                  lfc2 = ifelse(diff_Tray_CageCage3 == 1, 
                                round(lfc_Tray_CageCage3, 2), 0),
                  lfc3 = ifelse(diff_Tray_CageCage3_Tray_CageCage2 == 1, 
                                round(lfc_Tray_CageCage3_Tray_CageCage2, 2), 0)) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "value") %>%
    dplyr::arrange(taxon)

df_fig_pair2_CC = res_pair_CC %>%
    dplyr::filter(diff_Tray_CageCage2 == 1 |
                      diff_Tray_CageCage3 == 1 | 
                      diff_Tray_CageCage3_Tray_CageCage2 == 1) %>%
    dplyr::mutate(lfc1 = ifelse(passed_ss_Tray_CageCage2 == 1 & diff_Tray_CageCage2 == 1, 
                                "#00A087FF", "black"),
                  lfc2 = ifelse(passed_ss_Tray_CageCage3 == 1 & diff_Tray_CageCage3 == 1, 
                                "#00A087FF", "black"),
                  lfc3 = ifelse(passed_ss_Tray_CageCage3_Tray_CageCage2 == 1 & diff_Tray_CageCage3_Tray_CageCage2 == 1, 
                                "#00A087FF", "black")) %>%
    tidyr::pivot_longer(cols = lfc1:lfc3, 
                        names_to = "group", values_to = "color") %>%
    dplyr::arrange(taxon)

df_fig_pair_CC = df_fig_pair1_CC %>%
    dplyr::left_join(df_fig_pair2_CC, by = c("taxon", "group"))

df_fig_pair_CC$group = recode(df_fig_pair_CC$group, 
                          `lfc1` = "cage_2 - cage_1",
                          `lfc2` = "cage_3 - cage_1",
                          `lfc3` = "cage_3 - cage_2")

df_fig_pair_CC$group = factor(df_fig_pair_CC$group, 
                          levels = c("cage_2 - cage_1",
                                     "cage_3 - cage_1", 
                                     "cage_3 - cage_2"))


fig_pair_CC = df_fig_pair_CC %>%
    ggplot(aes(x = group, y = taxon, fill = value)) + 
    geom_tile(color = "black") +
    scale_fill_gradient2(low = "#002F6C", high = "#AC0123", mid = "white", 
                         na.value = "white", midpoint = 0, limit = c(-4, 4),
                         name = NULL) +
    geom_text(aes(group, taxon, label = value, color = color), size = 4) +
    scale_color_identity(guide = FALSE) +
    labs(x = NULL, y = NULL, title = "cage_insC") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
fig_pair_CC

```




```{r}

multipplot <- p3a <- ggpubr::ggarrange(fig_pair_AT, fig_pair_AC, fig_pair_BT, fig_pair_BC, fig_pair_CT, fig_pair_CC , 
                                       ncol = 2, 
                                       nrow = 3, 
                                       align = "v",
                                       widths = c(1,1))

multipplot

```




