---
title: "Ins_comp_3e-alpha_diversity_metrics"
author: "laura"
date: "2024-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The project

Does rearing in a different insectary affect the mosquito microbiome?

# The setup

one cohort of Ae aegypti eggs were split into 3 and reared in standard conditions to see how their microbiomes varied between insectaries at diff life stages.


# Install packages

```{r}

library(phyloseq)
library(ggplot2)
library(dplyr) 
library(MicrobiotaProcess)
library(gghalves)
library(ggh4x)
library(ggpubr)
library(dunn.test)
library(scales) 
library(pairwiseAdonis)

```

# import data

```{r}
ps = readRDS("physeq_obj_3974") # currently phyloseq obj

```


```{r}
# convert phyloseq object to mpse to use with MicrobiotaProcess package
mpse <- ps %>% as.MPSE() 

mpse

```

## alpha diversity

First do the calculations

```{r}
mpse %<>% 
    mp_cal_alpha(.abundance=Abundance)
mpse

```


```{r}
temp <- mp_extract_sample(mpse)

```

```{r}
colours <- c(
  "AdultFemale" = "#56187D", 
  "FishFood" = "#AC0123", 
  "Larva" = "#006347", 
  "LarvalWater" = "#97D0A7", 
  "Sugar" = "#EF9F26", 
  "Sugar2" = "#E6CD69", 
  "Water" = "#7AA8D7"
)


colours2 <- c(
  "C-ACC" = "#00A087FF",
  "A-CTID" = "#3C5488FF",
  "B-MK" = "#9C58A1"
)

```


### alpha div plot

```{r}

plot_alpha <- ggplot(temp, aes(x=Building, y=Shannon, fill=Building)) +
   geom_boxplot(outlier.shape = NA) +
  geom_point(position=position_jitterdodge(jitter.width=2, dodge.width = 0), pch=21, aes(x=Building, y=Shannon, fill=Building)) + 
   scale_fill_manual(values = colours2) +
  facet_grid(~factor(Sample_Type2, levels=c("Water", "FishFood", "LarvalWater", "Larva", "Sugar", "AdultFemale")), scales = "free", space = "free")  +
  xlab("") +
  ylab("Shannon's Index")  +
  theme(axis.text.x=element_blank()) +
  theme_bw()
  
plot_alpha

```


## alpha stats

### Are there differences between sample types?

```{r}

#Use kruskal wallis, a non-parametric test for global statistic 
kruskal.test(Shannon ~ Sample_Type2, data = temp)

```

Yes, so where are the differences?


```{r}
dunn.test(temp$Shannon, temp$Sample_Type2, method = "bonferroni") 

```
There are many differences here

### now, within each sample type, are there differences between insectaries?

Here, I am subsetting the data according to sample type then performing kruskal wallace test within each sample type to determine whether there are differences between insectary (called 'building' in metadata).

adult female

```{r}

shannon_adult <- subset(temp, temp$Sample_Type2 == "AdultFemale")

kruskal.test(Shannon ~ Building, data = shannon_adult) # p-value = 0.3824 - no difference


```

Larvae

```{r}

shannon_larvae <- subset(temp, temp$Sample_Type2 == "Larva")

kruskal.test(Shannon ~ Building, data = shannon_larvae) # p-value = 2.099e-10 - difference


```

```{r}

dunn.test(shannon_larvae$Shannon, shannon_larvae$Building, method = "bonferroni") 

```

Larval Water

```{r}

shannon_larval_water <- subset(temp, temp$Sample_Type2 == "LarvalWater")

kruskal.test(Shannon ~ Building, data = shannon_larval_water) # p-value = 1.44e-05 - difference

```

```{r}

dunn.test(shannon_larval_water$Shannon, shannon_larval_water$Building, method = "bonferroni") 

```


Water

```{r}

shannon_water <- subset(temp, temp$Sample_Type2 == "Water")

kruskal.test(Shannon ~ Building, data = shannon_water) # p-value = 0.2921 - no difference

```

Sugar

```{r}

shannon_sugar <- subset(temp, temp$Sample_Type2 == "Sugar")

kruskal.test(Shannon ~ Building, data = shannon_sugar) # p-value = 0.003042 - difference

```

```{r}

dunn.test(shannon_sugar$Shannon, shannon_sugar$Building, method = "bonferroni") 


```

fish food

```{r}

shannon_food <- subset(temp, temp$Sample_Type2 == "FishFood")

kruskal.test(Shannon ~ Building, data = shannon_food) # p-value = 0.1133 - no difference

```

